---
title: "UCLA Library Data Science Center"
subtitle: "Annual Service Report"
date: today
format:
  html:
    embed-resources: true
    toc: false
    css: styles-report.css
    theme: flatly
params:
  report_year: 2024
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(knitr)

report_year   <- params$report_year
prev_year     <- report_year - 1

summary   <- read_csv("data/processed/consultations/consultation_audit_2023_2025_summary_tagged.csv",
                      show_col_types = FALSE)
by_year   <- read_csv("data/processed/consultations/consultation_audit_2023_2025_by_year_group.csv",
                      show_col_types = FALSE)
coverage  <- read_csv("data/processed/consultations/consultation_audit_source_coverage.csv",
                      show_col_types = FALSE)
highlights <- read_csv("data/processed/consultations/qualitative_highlights.csv",
                       show_col_types = FALSE)
case_studies <- read_csv("data/processed/consultations/impact_case_studies.csv",
                         show_col_types = FALSE)

# Helper: pull a single KPI value by metric_id
kpi <- function(id) {
  v <- summary |> filter(metric_id == id) |> pull(value)
  if (length(v) == 0) return("—")
  as.character(v)
}

# Year-filtered helpers
consults_year <- function(yr) {
  by_year |> filter(year == yr) |> pull(n) |> sum()
}
```

---

## Executive Summary

This report summarizes service activity and infrastructure for the UCLA Library Data
Science Center (DSC) for `r report_year`. DSC provides consulting, instruction, and
research infrastructure support to UCLA researchers, students, and staff.

::: {.callout-note}
**Data coverage note:** Figures reflect available source data through
`r format(max(as.Date(coverage$max_date)), "%B %Y")`.
Full methodology and counting rules are in the
[Metrics Governance document](https://ucla-data-science-center.github.io/dsc-stats-reports/metrics-governance.html).
:::

### At a Glance — `r report_year`

```{r kpi-table}
# Filter by_year for report_year; compute DSC and DataSquad splits
yr_data <- by_year |> filter(year == report_year)
dsc_n   <- yr_data |> filter(group == "DSC")    |> pull(n) |> sum()
dsq_n   <- yr_data |> filter(group == "Datasquad") |> pull(n) |> sum()
total_n <- yr_data |> pull(n) |> sum()

# TODO: wire in workshop attendance from instruction data when available
# TODO: wire in Dataverse download count from infrastructure data

tribble(
  ~Metric,                              ~Value,
  "Direct consultations (total)",       format(total_n, big.mark = ","),
  "  — DSC staff consultations",        format(dsc_n,   big.mark = ","),
  "  — DataSquad consultations",        format(dsq_n,   big.mark = ","),
  "Workshop attendees",                 "— (TODO: wire in instruction data)",
  "Dataverse dataset downloads",        "— (TODO: wire in infrastructure data)",
  "AWS infrastructure spend",           "— (TODO: wire in cost data)"
) |>
  kable(col.names = c("Metric", report_year), align = c("l", "r"))
```

---

## Direct Service

DSC and DataSquad staff provided direct consultations to UCLA researchers across
schools and departments. Consultations include scheduled appointments tracked in
LibInsight and walk-in sessions at the DataSquad desk.

```{r consults-yoy}
# Year-over-year comparison
yoy <- by_year |>
  filter(year %in% c(prev_year, report_year)) |>
  group_by(year) |>
  summarise(total = sum(n), .groups = "drop") |>
  pivot_wider(names_from = year, values_from = total) |>
  mutate(change = .data[[as.character(report_year)]] - .data[[as.character(prev_year)]],
         pct    = round(change / .data[[as.character(prev_year)]] * 100, 1))

by_year |>
  filter(year %in% c(prev_year, report_year)) |>
  pivot_wider(names_from = year, values_from = n, values_fill = 0) |>
  rename(Group = group) |>
  kable(align = c("l", "r", "r"),
        caption = paste("Consultations by service area:", prev_year, "vs", report_year))
```

*Year-over-year change: `r if(nrow(yoy) > 0 && !is.na(yoy$change[1])) paste0(ifelse(yoy$change[1] >= 0, "+", ""), yoy$change[1], " (", yoy$pct[1], "%)") else "—"`*

---

## DataSquad Project Work

Beyond direct consultations, DataSquad student consultants provide sustained project
support tracked via Trello task cards. Task cards represent assigned work units and
are not equivalent to consultations.

```{r task-cards}
# TODO: expand when GitHub Projects data is integrated
tribble(
  ~Metric,                                   ~Value,
  "Reportable task cards (all years)",       kpi("datasquad_trello_task_cards_reportable_count"),
  "Comment-weighted task activity (reportable)", kpi("datasquad_trello_task_activity_comment_weighted_reportable_count")
) |>
  kable(col.names = c("Metric", "Count"), align = c("l", "r"))
```

*Note: Task card data reflects Trello (2021–2024). GitHub Projects integration
is planned for 2025 onwards.*

---

## Research Infrastructure

```{r infrastructure-placeholder}
# TODO: pull from infrastructure data files once refreshed
# Key metrics to wire in:
#   - Dataverse: total datasets, files, cumulative downloads (datasets_files_published_monthly.csv)
#   - AWS: annual total cost (aws_costs_all_normalized.csv)
#   - Potree: current S3 storage size (live CloudWatch / AWS MCP)
```

*Infrastructure metrics (Dataverse growth, AWS costs, S3 storage) to be wired in
after data refresh. See [Infrastructure page](https://ucla-data-science-center.github.io/dsc-stats-reports/infrastructure.html) for current figures.*

---

## Impact Highlights

```{r case-studies}
active_cases <- case_studies |>
  filter(include_in_dashboard == TRUE) |>
  arrange(display_order)

for (i in seq_len(nrow(active_cases))) {
  row <- active_cases[i, ]
  cat(paste0("### ", row$case_title, "\n\n"))
  cat(paste0("**Partners:** ", row$partners, "\n\n"))
  cat(paste0(row$summary, "\n\n"))
  if (!is.na(row$measurable_outcomes) && nchar(row$measurable_outcomes) > 0) {
    cat(paste0("**Outcomes:** ", row$measurable_outcomes, "\n\n"))
  }
  if (!is.na(row$impact_signals) && nchar(row$impact_signals) > 0) {
    cat(paste0("**Impact:** ", row$impact_signals, "\n\n"))
  }
  cat("---\n\n")
}
```

```{r highlights}
active_highlights <- highlights |>
  filter(include_in_dashboard == TRUE)

if (nrow(active_highlights) > 0) {
  cat("### Selected Service Highlights\n\n")
  for (i in seq_len(nrow(active_highlights))) {
    row <- active_highlights[i, ]
    cat(paste0("- **", row$period_label, " — ", row$request_theme, ":** ",
               row$outcome, "\n"))
  }
}
```

---

## About This Report

This report is generated from the same data pipeline that powers the
[DSC Statistics Dashboard](https://ucla-data-science-center.github.io/dsc-stats-reports/).
All figures reflect data baked in at render time — this is not a live feed.

For methodology, counting rules, and data limitations see the
[About & methodology page](https://ucla-data-science-center.github.io/dsc-stats-reports/about.html).

*Rendered: `r format(Sys.Date(), "%B %d, %Y")`*
