---
title: "Consulting and project support"
format:
  html:
    toc: true
    code-fold: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
source("src/etl/suppression.R")

# ── Consultation records: public aggregates only ───────────
# The row-level records file is not in the public repository (it holds free-text
# fields that can contain personal contact details). These suppressed aggregates
# are built locally by src/etl/consult_public_aggregates.R.
pub <- function(name) {
  path <- file.path("data/processed/consultations/public", name)
  if (file.exists(path)) read_csv(path, show_col_types = FALSE) else tibble()
}
rec_year   <- pub("consult_records_by_year.csv")
rec_dept   <- pub("consult_records_by_department.csv")
rec_affil  <- pub("consult_records_by_affiliation.csv")
rec_team   <- pub("consult_records_by_team.csv")
rec_year_team <- pub("consult_records_by_year_team.csv")
rec_window <- pub("consult_records_window.csv")
data_loaded <- nrow(rec_year) > 0
rec_window_label <- if (nrow(rec_window) > 0) {
  paste(format(rec_window$first_date, "%b %Y"), "to", format(rec_window$last_date, "%b %Y"))
} else "window unknown"
suppressed_note <- function(n, unit) {
  if (length(n) == 0 || is.na(n) || n == 0) "" else
    sprintf(" %d %s with fewer than 5 are combined into one row, or omitted if that row is also under 5.", n, unit)
}

# ── Audit outputs (all graceful) ──────────────────────────────
read_if <- function(path) {
  if (file.exists(path)) read_csv(path, show_col_types = FALSE) else tibble()
}

audit_summary   <- read_if("data/processed/consultations/consultation_audit_2023_2025_summary_tagged.csv")
audit_task_mode <- read_if("data/processed/consultations/consultation_audit_2023_2025_by_year_task_mode.csv")
audit_coverage  <- read_if("data/processed/consultations/consultation_audit_source_coverage.csv")
by_year_src     <- read_if("data/processed/consultations/consultation_audit_2023_2025_by_year_source.csv")
qual_highlights      <- read_if("data/processed/consultations/qualitative_highlights.csv")
impact_case_studies  <- read_if("data/processed/consultations/impact_case_studies.csv")

kpi <- function(key) {
  if (nrow(audit_summary) == 0) return("—")
  x <- audit_summary |> filter(metric_id == key) |> pull(value)
  if (length(x) == 0) "—" else comma(x[[1]])
}

no_data_msg <- function() {
  cat("*Data file not loaded. Run the ETL scripts to populate this chart.*\n")
}

# ── Pre-computed alt text for charts ─────────────────────────
if (data_loaded) {
  .yt <- rec_year_team |> mutate(group = recode(group, Datasquad = "walk-in sign-ins", DSC = "scheduled appointments", Other = "other"))
  alt_year <- sprintf(
    "Horizontal stacked bar chart: consultation records file by year and record type, %s. %s. 2024 covers January to early June only.",
    rec_window_label,
    paste(sprintf("%d: %s", .yt$year, paste(.yt$group, .yt$records)), collapse = "; ")
  )
  alt_group <- sprintf(
    "Bar chart: consultation records by team, %s. %s.",
    rec_window_label, paste(rec_team$group, rec_team$records, sep = " = ", collapse = "; ")
  )
} else {
  alt_year  <- "Consultation records by year (data not loaded)"
  alt_group <- "Consultation records by team (data not loaded)"
}

data_range_label <- if (nrow(audit_coverage) > 0) {
  yr_min <- format(min(as.Date(audit_coverage$min_date)), "%Y")
  yr_max <- format(max(as.Date(audit_coverage$max_date)), "%Y")
  if (yr_min == yr_max) yr_min else paste0(yr_min, "\u2013", yr_max)
} else {
  "All Time"
}

# YoY comparison: 2023 → 2024 (two complete calendar years in this audit)
# 2025 data is partial (libinsights export ends March 2025) — excluded from YoY
yoy_fmt <- function(n_recent, n_prior, yr_recent = 2024, yr_prior = 2023) {
  if (!length(n_recent) || !length(n_prior) || is.na(n_recent) || is.na(n_prior) || n_prior == 0) return("")
  pct   <- round((n_recent - n_prior) / n_prior * 100)
  arrow <- if (pct > 0) "\u2191" else if (pct < 0) "\u2193" else "\u2192"
  sprintf("%s %+d%% \u00b7 %s in %d vs %s in %d",
          arrow, pct, comma(n_recent), yr_recent, comma(n_prior), yr_prior)
}

yoy_direct <- local({
  if (nrow(by_year_src) == 0) return("")
  d   <- by_year_src |> filter(source != "trello") |>
         group_by(year) |> summarise(n = sum(n), .groups = "drop")
  yoy_fmt(d$n[d$year == 2024], d$n[d$year == 2023])
})

yoy_tasks <- local({
  if (nrow(audit_task_mode) == 0) return("")
  d <- audit_task_mode |> filter(task_metric == "trello_reportable_task_cards")
  yoy_fmt(d$n[d$year == 2024], d$n[d$year == 2023])
})
```

*How much consulting DSC recorded, who used it, and what project work followed. Definitions and the separate audit are at the end.*

```{r series_setup, include=FALSE}
consult_series <- read_if("data/processed/canonical/consultations_by_year.csv")
series_total <- read_if("data/processed/canonical/headline_aggregates.csv") |>
  filter(metric_id == "consultations_recorded_series") |> pull(display)
if (length(series_total) == 0) series_total <- "—"
last_full_year <- if (nrow(consult_series) > 0) max(consult_series$year[!consult_series$is_floor]) else NA
```

DSC recorded **`r series_total` scheduled research consultations** from 2017 to September 2026 (claim-register #10), with DSC staff and DataSquad student consultants alike. These are appointments, not people: one researcher can book many.

::: {.callout-note collapse="true" title="Definition and sources"}
The series counts scheduled appointments from Calendly (2017 to mid-2024), then LibInsight and LibCal. DataSquad's in-person walk-in sign-ins are not included. Internal meetings and canceled appointments are removed.
:::

## Recorded consultations over time {#recorded-consultations-since-2017}

*Scheduled appointments by calendar year, 2017 to September 2026.*

```{r series_chart, echo=FALSE, fig.height=4.2, fig.alt=if (nrow(consult_series) > 0) paste0("Horizontal bar chart of recorded scheduled consultations by year, earliest at top. Before the recording gap: ", paste(sprintf("%d, %s", consult_series$year[!consult_series$is_floor], consult_series$display[!consult_series$is_floor]), collapse = "; "), ". Incomplete records after mid-2024: ", paste(sprintf("%d, %s", consult_series$year[consult_series$is_floor], consult_series$display[consult_series$is_floor]), collapse = "; "), ". 2026 is partial.") else "Consultation series not loaded"}
if (nrow(consult_series) > 0) {
  consult_series |>
    mutate(status = if_else(is_floor, "Incomplete records, 2024-2026", "Before the recording gap"),
           status = factor(status, levels = c("Before the recording gap", "Incomplete records, 2024-2026")),
           label = if_else(year == max(year), paste0(display, " (partial year)"), display),
           year = factor(year, levels = rev(sort(unique(year))))) |>
    ggplot(aes(x = value, y = year, fill = status)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = label), hjust = -0.1, size = 3.5) +
    facet_grid(status ~ ., scales = "free_y", space = "free_y") +
    scale_fill_manual(values = c("Before the recording gap" = "#2774AE",
                                 "Incomplete records, 2024-2026" = "#8C8279"), guide = "none") +
    scale_x_continuous(expand = expansion(mult = c(0, 0.25))) +
    labs(x = "Recorded consultations", y = NULL) +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
          strip.text.y = element_text(angle = 0, hjust = 0, face = "bold"))
}
```

**Records are incomplete after mid-2024.** Calendly ended, LibCal wasn't in use until May 2025, and two reorganizations and staff departures cut logging. The later counts are what was recorded, not estimates of the missing activity, so this series cannot show how demand changed after `r last_full_year`.

::: {.callout-note collapse="true" title="View annual counts"}
```{r series_table, echo=FALSE}
if (nrow(consult_series) > 0) {
  consult_series |>
    transmute(Year = as.character(year), `Recorded consultations` = display,
              Records = if_else(is_floor, "Incomplete after mid-2024", "Before the recording gap"),
              Source = source) |>
    knitr::kable(align = c("l", "r", "l", "l"))
}
```
:::

## Who used consulting? {#direct-consultations}

*January 2021 to June 2024 records with a recorded department. Consultation records, not unique people.*

```{r dept_setup, include=FALSE}
other_label <- "Other departments (each fewer than 5)"
dept_named <- rec_dept |> filter(department != other_label) |> arrange(desc(records), department)
dept_other <- rec_dept |> filter(department == other_label) |> pull(records)
dept_top <- head(dept_named, 10)
alt_dept <- if (nrow(dept_top) > 0) paste0("Horizontal bar chart of the ten departments with the most consultation records, January 2021 to June 2024, largest first: ",
  paste(dept_top$department, dept_top$records, collapse = "; "), ".") else "Department chart (data not loaded)"
```

```{r chart_dept, echo=FALSE, fig.height=4.8, fig.alt=alt_dept}
if (data_loaded && nrow(dept_top) > 0) {
  dept_top |>
    mutate(department = factor(str_wrap(department, 28), levels = rev(str_wrap(department, 28)))) |>
    ggplot(aes(x = records, y = department)) +
    geom_col(fill = "#2774AE", width = 0.7) +
    geom_text(aes(label = records), hjust = -0.2, size = 3.5) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
    labs(x = "Consultation records", y = NULL, title = "Ten largest recorded departments") +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank())
} else { no_data_msg() }
```

`r if (length(dept_other)) sprintf("Another %s records come from departments with fewer than 5 each; they are grouped in the table below, not shown as one bar.", comma(dept_other)) else ""`

::: {.callout-note collapse="true" title="All departments (searchable)"}
```{r dept_table, echo=FALSE}
if (data_loaded) {
  bind_rows(dept_named, rec_dept |> filter(department == other_label)) |>
    DT::datatable(rownames = FALSE, colnames = c("Department", "Consultation records"),
                  options = list(pageLength = 15, dom = "ftp"),
                  caption = paste0("Consultation records file, ", rec_window_label, ". Named departments first; the combined group is listed last.",
                                   suppressed_note(rec_window$suppressed_departments, "departments")))
}
```
:::

### Consultations by researcher role

```{r affil_setup, include=FALSE}
total_records <- sum(rec_year$records)
affil_listed <- sum(rec_affil$records)
affil_missing <- total_records - affil_listed
# The difference is "no role recorded" only if every role record is dated (true as of
# 2026-09-29: the one undated record has no role) and the combined small-role row is
# published. Otherwise fall back to the neutral wording.
small_role_row_shown <- any(grepl("^Other affiliations", rec_affil$ucla_affiliation)) ||
  coalesce(rec_window$suppressed_affiliations[1], 0) == 0
missing_role_note <- if (small_role_row_shown && affil_missing >= 0) {
  sprintf("%s of the %s records have no role recorded and are not shown. That is missing information, separate from the small roles combined into one row.",
          comma(affil_missing), comma(total_records))
} else {
  sprintf("%s of the %s records are not represented in this breakdown.", comma(max(affil_missing, 0)), comma(total_records))
}
alt_affil <- if (nrow(rec_affil) > 0) paste0("Horizontal bar chart of consultation records by researcher role, ", rec_window_label, ": ",
  paste(rec_affil$ucla_affiliation, rec_affil$records, collapse = "; "), ".") else "Affiliation chart (data not loaded)"
```

```{r chart_affil, echo=FALSE, fig.height=3.2, fig.alt=alt_affil}
if (data_loaded && nrow(rec_affil) > 0) {
  rec_affil |>
    mutate(role = str_wrap(ucla_affiliation, 24),
           role = factor(role, levels = rev(role))) |>
    ggplot(aes(x = records, y = role)) +
    geom_col(fill = "#2774AE", width = 0.7) +
    geom_text(aes(label = records), hjust = -0.2, size = 3.5) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
    labs(x = "Consultation records", y = NULL) +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank())
} else { no_data_msg() }
```

Same records file, `r rec_window_label`. `r if (data_loaded) missing_role_note else ""`

## Student project work

```{r task_setup, include=FALSE}
task_cards <- kpi("datasquad_trello_task_cards_reportable_count")
```

DataSquad students documented **`r task_cards` project tasks** from 2023 to 2025. A task is one piece of follow-on project work tracked on the team's Trello board, such as cleaning a dataset, writing code, or preparing a map, after a consultation turned into a project. Tasks are not consultations and are not added to the consultation counts. Tracking on Trello became routine in 2023, so earlier project work is not counted.

## What this work supported {#research-impact}

Some consultations became multi-year collaborations. Two examples, with full accounts on the [Research enabled page](research.qmd#three-collaborations):

- [Police helicopter surveillance](research.qmd#police-helicopter-surveillance-carceral-ecologies): DSC staff and DataSquad students cleaned and joined a lab's flight data, and one data-loading step went from about 318 minutes to about 2.
- [Deaths in custody](research.qmd#deaths-in-custody-biocritical-studies-lab): DataSquad students merged and reconciled several sources into a national dataset the lab now maintains and updates itself.

## Sources and counting rules {#sources}

Three different record sets appear on this page. They are kept separate because they count different things:

| Record set | Window | Used for |
|---|---|---|
| Scheduled consultation series | 2017 to September 2026 | The headline count and the annual chart |
| Consultation records file | `r rec_window_label` | Department and researcher-role breakdowns (includes DataSquad walk-in sign-ins) |
| 2023-2025 consultation audit | January 2023 to 2025 | Checking the records; project task counts |

::: {.callout-note collapse="true" title="The 2023-2025 consultation audit"}
The audit combined DSC's LibInsight appointments with DataSquad's LibInsight appointments and walk-in sign-ins, excluded canceled appointments, and removed likely duplicate manual logs. It found **`r kpi("direct_consults_combined_deduped_count")` direct consultations** (claim-register #7). It is a check on the records, not a second headline: its definition and window differ from the series above.

- DSC staff appointments (LibInsight): `r kpi("dsc_direct_consults_libinsights_count")`
- DataSquad consultations (LibInsight and sign-ins): `r kpi("datasquad_direct_consults_count")`
- A narrower combined count using fewer DataSquad sources: `r kpi("direct_consults_dsc_plus_datasquad_count")`. The 741 total is larger because it includes more DataSquad sign-in sources; the parts above do not sum to it.
- Likely duplicate manual logs removed: about `r kpi("libinsight_manual_logs_heuristic_duplicate_count")` (same provider, same minute)

```{r coverage, echo=FALSE}
if (nrow(audit_coverage) > 0) {
  audit_coverage |>
    mutate(min_date = as.Date(min_date), max_date = as.Date(max_date)) |>
    arrange(source, subtype) |>
    transmute(
      Source        = source,
      `Record type` = subtype,
      Rows          = comma(rows),
      `Date range`  = paste(format(min_date, "%Y-%m-%d"), "–",
                            format(max_date, "%Y-%m-%d"))
    ) |>
    knitr::kable(caption = "Source coverage: rows and date ranges per data system")
}
```
:::

::: {.callout-note collapse="true" title="Task-tracking methods"}
- Documented project tasks are Trello cards on the DataSquad board, excluding the `Inbox`, `Future Projects/Ideas`, and `Team Info` lists and including archived completed cards. All cards, including those lists: `r kpi("datasquad_trello_task_cards_all_count")`.
- The audit also weighted each card by its comments (`r kpi("datasquad_trello_task_activity_comment_weighted_reportable_count")` activity rows). That measure is kept for internal checks and is not comparable to consultation counts.
- Growth in task counts after 2023 partly reflects wider use of Trello, not only more work.
- Full rules: [Metrics Governance](metrics-governance.md) · [About & methodology](about.qmd)
:::

---

**See also:** [Instruction <span aria-hidden="true">→</span>](instruction.qmd) workshop attendance · [Infrastructure <span aria-hidden="true">→</span>](infrastructure.qmd) Dataverse and AWS metrics · [About & methodology](about.qmd)

::: {.callout-note collapse="true" title="Cite this page"}
UCLA Library Data Science Center. (`r year(Sys.Date())`). *Consulting and project support*. Retrieved from <https://ucla-data-science-center.github.io/dsc-stats-reports/consulting.html>
:::

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
