---
title: "Consulting Statistics"
format:
  html:
    toc: true
    code-fold: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(plotly)
library(crosstalk)

# ── Consultation records (graceful fallback) ──────────────────
data_loaded <- FALSE
dsc_consult  <- tibble()

for (p in c(
  "data/processed/consultations/dsc_consult_merged.rds",
  "data/consultations/Data/dsc_consult_merged.rds",
  "data/consultations/dsc_consult_merged.rds"
)) {
  if (file.exists(p)) {
    dsc_consult <- readRDS(p) |>
      mutate(ucla_affiliation = case_when(
        ucla_affiliation %in% c("Graduate", "Graduate Student",
                                "Visiting Graduate Student") ~ "Graduate Student",
        ucla_affiliation %in% c("Undergrauate 3rd & Undergraduate",
                                "Undergraduate") ~ "Undergraduate",
        TRUE ~ ucla_affiliation
      ))
    data_loaded <- TRUE
    break
  }
}

if (!data_loaded) {
  message("dsc_consult_merged.rds not found — charts will show placeholder text.")
}

# ── Audit outputs (all graceful) ──────────────────────────────
read_if <- function(path) {
  if (file.exists(path)) read_csv(path, show_col_types = FALSE) else tibble()
}

audit_summary   <- read_if("data/processed/consultations/consultation_audit_2023_2025_summary_tagged.csv")
audit_task_mode <- read_if("data/processed/consultations/consultation_audit_2023_2025_by_year_task_mode.csv")
audit_coverage  <- read_if("data/processed/consultations/consultation_audit_source_coverage.csv")
by_year_src     <- read_if("data/processed/consultations/consultation_audit_2023_2025_by_year_source.csv")
qual_highlights      <- read_if("data/processed/consultations/qualitative_highlights.csv")
impact_case_studies  <- read_if("data/processed/consultations/impact_case_studies.csv")

kpi <- function(key) {
  if (nrow(audit_summary) == 0) return("—")
  x <- audit_summary |> filter(metric_id == key) |> pull(value)
  if (length(x) == 0) "—" else comma(x[[1]])
}

no_data_msg <- function() {
  cat("*Data file not loaded. Run the ETL scripts to populate this chart.*\n")
}

# ── Pre-computed alt text for charts ─────────────────────────
if (data_loaded) {
  .yr <- dsc_consult |> filter(!is.na(start_date_time)) |>
    count(year = year(start_date_time)) |> arrange(year)
  alt_year <- sprintf(
    "Horizontal bar chart: DSC consultations by year from %d to %d. %s.",
    min(.yr$year), max(.yr$year),
    paste(.yr$year, .yr$n, sep = ": ", collapse = "; ")
  )
  .dept <- dsc_consult |>
    filter(!is.na(department), !department %in% c("DSC", "Library", "UCLA")) |>
    count(department, sort = TRUE) |> head(15)
  alt_dept <- sprintf(
    "Horizontal bar chart: top %d departments by consultation count. Highest: %s (%d).",
    nrow(.dept), .dept$department[1], .dept$n[1]
  )
  .affil <- dsc_consult |> filter(!is.na(ucla_affiliation)) |>
    count(ucla_affiliation, sort = TRUE) |> head(6)
  alt_affil <- sprintf(
    "Horizontal bar chart: consultations by researcher affiliation. Top group: %s (%d).",
    .affil$ucla_affiliation[1], .affil$n[1]
  )
  .grp <- dsc_consult |> filter(!is.na(group)) |> count(group)
  alt_group <- sprintf(
    "Bar chart: consultations by team. %s.",
    paste(.grp$group, .grp$n, sep = " = ", collapse = "; ")
  )
} else {
  alt_year  <- "Consultations by year (data not loaded)"
  alt_dept  <- "Top departments by consultations (data not loaded)"
  alt_affil <- "Consultations by affiliation (data not loaded)"
  alt_group <- "DSC vs DataSquad split (data not loaded)"
}

data_range_label <- if (nrow(audit_coverage) > 0) {
  yr_min <- format(min(as.Date(audit_coverage$min_date)), "%Y")
  yr_max <- format(max(as.Date(audit_coverage$max_date)), "%Y")
  if (yr_min == yr_max) yr_min else paste0(yr_min, "\u2013", yr_max)
} else {
  "All Time"
}

# YoY comparison: 2023 → 2024 (two complete calendar years in this audit)
# 2025 data is partial (libinsights export ends March 2025) — excluded from YoY
yoy_fmt <- function(n_recent, n_prior, yr_recent = 2024, yr_prior = 2023) {
  if (!length(n_recent) || !length(n_prior) || is.na(n_recent) || is.na(n_prior) || n_prior == 0) return("")
  pct   <- round((n_recent - n_prior) / n_prior * 100)
  arrow <- if (pct > 0) "\u2191" else if (pct < 0) "\u2193" else "\u2192"
  sprintf("%s %+d%% \u00b7 %s in %d vs %s in %d",
          arrow, pct, comma(n_recent), yr_recent, comma(n_prior), yr_prior)
}

yoy_direct <- local({
  if (nrow(by_year_src) == 0) return("")
  d   <- by_year_src |> filter(source != "trello") |>
         group_by(year) |> summarise(n = sum(n), .groups = "drop")
  yoy_fmt(d$n[d$year == 2024], d$n[d$year == 2023])
})

yoy_tasks <- local({
  if (nrow(audit_task_mode) == 0) return("")
  d <- audit_task_mode |> filter(task_metric == "trello_reportable_task_cards")
  yoy_fmt(d$n[d$year == 2024], d$n[d$year == 2023])
})
```

*For DSC staff and library administration — direct consultation counts, project task metrics, and patron breakdowns.*

## At a Glance (`r data_range_label`)

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*

::: {.kpi-row .grid}

::: {.g-col-6 .g-col-md-3 .kpi-card}
[Direct Consultations]{.kpi-label}
[`r kpi("direct_consults_combined_deduped_count")`]{.kpi-value}
[DSC + DataSquad combined; canceled excluded`r if (nchar(yoy_direct) > 0) paste0(" \u00b7 ", yoy_direct)`]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[DSC Staff]{.kpi-label}
[`r kpi("dsc_direct_consults_libinsights_count")`]{.kpi-value}
[LibInsight scheduled exports]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[DataSquad]{.kpi-label}
[`r kpi("datasquad_direct_consults_count")`]{.kpi-value}
[Walk-in sign-ins]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[Task Cards (Reportable)]{.kpi-label}
[`r kpi("datasquad_trello_task_cards_reportable_count")`]{.kpi-value}
[Trello; backlog/admin lists excluded`r if (nchar(yoy_tasks) > 0) paste0(" \u00b7 ", yoy_tasks)`]{.kpi-note}
:::

:::

## Direct Consultations

::: {.panel-tabset}

### By Year

```{r chart_year, echo=FALSE, fig.alt=alt_year}
if (data_loaded) {
  dsc_consult |>
    filter(!is.na(start_date_time)) |>
    count(year = year(start_date_time)) |>
    ggplot(aes(x = factor(year), y = n)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    labs(x = NULL, y = "Consultations") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

### By Department

```{r chart_dept, echo=FALSE}
if (data_loaded) {
  internal_depts <- c("DSC", "Library", "UCLA")
  dept_counts <- dsc_consult |>
    filter(!is.na(department), !department %in% internal_depts) |>
    count(department, sort = TRUE) |>
    rename(Department = department, Consultations = n)

  sd_dept <- SharedData$new(dept_counts)

  # Filter widget + linked chart + linked table
  filter_select("dept_filter", "Search department:", sd_dept, ~Department,
                multiple = FALSE)

  bscols(
    plot_ly(sd_dept, x = ~Consultations, y = ~reorder(Department, Consultations),
            type = "bar", orientation = "h",
            marker = list(color = "#2774AE"),
            hovertemplate = "<b>%{y}</b><br>Consultations: %{x}<extra></extra>") |>
      layout(yaxis = list(title = ""),
             xaxis = list(title = "Consultations"),
             margin = list(l = 5)),
    DT::datatable(sd_dept, rownames = FALSE,
                  colnames = c("Department", "Consultations"),
                  options = list(pageLength = 10, dom = "tp"),
                  caption = "All departments — linked to chart above")
  )
} else { no_data_msg() }
```

### By Affiliation

```{r chart_affil, echo=FALSE}
if (data_loaded) {
  affil_counts <- dsc_consult |>
    filter(!is.na(ucla_affiliation)) |>
    count(ucla_affiliation, sort = TRUE) |>
    filter(n >= 3) |>   # drop obvious data-entry noise (phone numbers, names)
    rename(Affiliation = ucla_affiliation, Consultations = n)

  sd_affil <- SharedData$new(affil_counts)

  filter_checkbox("affil_filter", "Filter affiliation:", sd_affil, ~Affiliation)

  bscols(
    plot_ly(sd_affil, x = ~Consultations, y = ~reorder(Affiliation, Consultations),
            type = "bar", orientation = "h",
            marker = list(color = "#2774AE"),
            hovertemplate = "<b>%{y}</b><br>Consultations: %{x}<extra></extra>") |>
      layout(yaxis = list(title = ""),
             xaxis = list(title = "Consultations"),
             margin = list(l = 5)),
    DT::datatable(sd_affil, rownames = FALSE,
                  colnames = c("Affiliation", "Consultations"),
                  options = list(pageLength = 10, dom = "t"),
                  caption = "Linked to chart above")
  )
} else { no_data_msg() }
```

### By Team

```{r chart_group, echo=FALSE, fig.alt=alt_group}
if (data_loaded) {
  dsc_consult |>
    filter(!is.na(group)) |>
    count(group) |>
    ggplot(aes(x = group, y = n, fill = group)) +
    geom_col() +
    scale_fill_manual(values = c("DSC" = "#2774AE", "Datasquad" = "#FFD100")) +
    labs(x = NULL, y = "Consultations") +
    theme_minimal() +
    theme(legend.position = "none")
} else { no_data_msg() }
```

:::

## Project Work & Task Metrics

These metrics represent follow-on project work assigned to DataSquad members. They are **not** direct consultations.

::: {.panel-tabset}

### Summary

```{r task_table, echo=FALSE}
if (nrow(audit_summary) == 0) {
  cat("*Run `Rscript src/etl/consultation_audit_summary.R` to generate task metrics.*\n")
} else {
  audit_summary |>
    filter(metric_id %in% c(
      "direct_consults_combined_deduped_count",
      "datasquad_trello_task_cards_reportable_count",
      "datasquad_trello_task_activity_comment_weighted_reportable_count",
      "datasquad_total_with_trello_reportable_count"
    )) |>
    transmute(
      Metric = recode(metric_id,
        direct_consults_combined_deduped_count =
          "Direct Consultations (Combined, Deduped)",
        datasquad_trello_task_cards_reportable_count =
          "DataSquad Task Cards (Reportable)",
        datasquad_trello_task_activity_comment_weighted_reportable_count =
          "DataSquad Task Activity (Reportable, Comment-Weighted)",
        datasquad_total_with_trello_reportable_count =
          "DataSquad Direct + Reportable Task Activity"
      ),
      Value = comma(value)
    ) |>
    knitr::kable(align = c("l", "r"))
}
```

### Task Activity by Year

```{r task_year, echo=FALSE, fig.alt="Grouped bar chart showing Trello task card counts and comment-weighted activity by year, split by all cards vs reportable-filtered cards."}
if (nrow(audit_task_mode) == 0) {
  cat("*Task-mode audit file not found.*\n")
} else {
  audit_task_mode |>
    filter(task_metric %in% c(
      "trello_all_task_cards", "trello_reportable_task_cards",
      "trello_all_comment_weighted", "trello_reportable_comment_weighted"
    )) |>
    mutate(
      task_metric = recode(task_metric,
        trello_all_comment_weighted        = "Activity (all)",
        trello_reportable_comment_weighted = "Activity (reportable)",
        trello_all_task_cards              = "Task cards (all)",
        trello_reportable_task_cards       = "Task cards (reportable)"
      ),
      metric_family = if_else(
        grepl("cards", task_metric), "Task cards (count)", "Task activity (comment-weighted)"
      )
    ) |>
    ggplot(aes(x = factor(year), y = n, fill = task_metric)) +
    geom_col(position = "dodge") +
    facet_wrap(~metric_family, scales = "free_y") +
    labs(
      x = "Year", y = "Count", fill = NULL,
      caption = "Trello historical. Reportable metric excludes backlog/admin/archived lists."
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")
}
```

:::

## Research Impact

These examples show how DSC and DataSquad support extends into sustained technical collaboration beyond single consultations.

```{r impact, echo=FALSE, results='asis'}
if (nrow(impact_case_studies) == 0) {
  cat("*Add rows to `data/processed/consultations/impact_case_studies.csv` to render case studies here.*\n")
} else {
  impact_case_studies |>
    filter(coalesce(include_in_dashboard, TRUE)) |>
    arrange(display_order) |>
    pwalk(function(case_title, partners, summary,
                   measurable_outcomes, impact_signals, ...) {
      cat("### ", case_title, "\n\n", sep = "")
      if (!is.na(partners) && partners != "")
        cat("**Partners:** ", partners, "\n\n", sep = "")
      if (!is.na(summary) && summary != "")
        cat(summary, "\n\n")
      if (!is.na(measurable_outcomes) && measurable_outcomes != "")
        cat("**Measurable outcomes:** ", measurable_outcomes, "\n\n", sep = "")
      if (!is.na(impact_signals) && impact_signals != "")
        cat("**Impact signals:** ", impact_signals, "\n\n", sep = "")
    })
}
```

```{r qualitative, echo=FALSE}
if (nrow(qual_highlights) > 0) {
  display_cols <- c("period_label", "workstream", "audience",
                    "request_theme", "work_performed", "outcome")
  qual_highlights |>
    select(all_of(intersect(display_cols, names(qual_highlights)))) |>
    rename_with(~str_replace_all(., "_", " ")) |>
    knitr::kable(align = "l")
}
```

---

**See also:** [Instruction <span aria-hidden="true">→</span>](instruction.qmd) workshop attendance · [Infrastructure <span aria-hidden="true">→</span>](infrastructure.qmd) Dataverse and AWS metrics · [About & methodology](about.qmd)

---

::: {.callout-note collapse="true" .data-notes title="Data notes"}
- **Direct consultations**: LibInsight scheduled exports + manual logs (DSC staff); DataSquad walk-in sign-in forms
- **Canceled appointments** excluded from all direct consultation counts
- **Deduplication**: manual LibInsight logs matched against scheduled exports using a same-provider/same-minute heuristic; ~`r kpi("libinsight_manual_logs_heuristic_duplicate_count")` duplicates removed
- **Task metrics** (Trello cards, comment-weighted activity) are shown separately — not comparable to consultation counts
- **Reportable filter**: excludes `Inbox`, `Future Projects/Ideas`, `Team Info` lists; includes archived completed cards
- **YoY comparisons** (2023 → 2024): direct consultation decline reflects shifting data sources and intake channels, not necessarily service volume; task card growth reflects expanded Trello adoption in 2024, not a proportional workload increase
- Full rules: [Metrics Governance](metrics-governance.md) · [About & methodology](about.qmd)
:::

```{r coverage, echo=FALSE}
if (nrow(audit_coverage) > 0) {
  audit_coverage |>
    mutate(min_date = as.Date(min_date), max_date = as.Date(max_date)) |>
    arrange(source, subtype) |>
    transmute(
      Source        = source,
      `Record type` = subtype,
      Rows          = comma(rows),
      `Date range`  = paste(format(min_date, "%Y-%m-%d"), "\u2013",
                            format(max_date, "%Y-%m-%d"))
    ) |>
    DT::datatable(
      rownames = FALSE,
      options  = list(pageLength = 10, dom = "tp"),
      caption  = "Source coverage: rows and date ranges per data system"
    )
}
```

::: {.callout-note collapse="true" title="Cite this page"}
UCLA Library Data Science Center. (`r year(Sys.Date())`). *Consulting Statistics*. Retrieved from <https://ucla-data-science-center.github.io/dsc-stats-reports/consulting.html>
:::

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
