---
title: "Consulting Statistics"
format:
  html:
    toc: true
    code-fold: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(plotly)
library(crosstalk)
source("src/etl/suppression.R")

# ── Consultation records: public aggregates only ───────────
# The row-level records file is not in the public repository (it holds free-text
# fields that can contain personal contact details). These suppressed aggregates
# are built locally by src/etl/consult_public_aggregates.R.
pub <- function(name) {
  path <- file.path("data/processed/consultations/public", name)
  if (file.exists(path)) read_csv(path, show_col_types = FALSE) else tibble()
}
rec_year   <- pub("consult_records_by_year.csv")
rec_dept   <- pub("consult_records_by_department.csv")
rec_affil  <- pub("consult_records_by_affiliation.csv")
rec_team   <- pub("consult_records_by_team.csv")
rec_year_team <- pub("consult_records_by_year_team.csv")
rec_window <- pub("consult_records_window.csv")
data_loaded <- nrow(rec_year) > 0
rec_window_label <- if (nrow(rec_window) > 0) {
  paste(format(rec_window$first_date, "%b %Y"), "to", format(rec_window$last_date, "%b %Y"))
} else "window unknown"
suppressed_note <- function(n, unit) {
  if (length(n) == 0 || is.na(n) || n == 0) "" else
    sprintf(" %d %s with fewer than 5 are combined into one row, or omitted if that row is also under 5.", n, unit)
}

# ── Audit outputs (all graceful) ──────────────────────────────
read_if <- function(path) {
  if (file.exists(path)) read_csv(path, show_col_types = FALSE) else tibble()
}

audit_summary   <- read_if("data/processed/consultations/consultation_audit_2023_2025_summary_tagged.csv")
audit_task_mode <- read_if("data/processed/consultations/consultation_audit_2023_2025_by_year_task_mode.csv")
audit_coverage  <- read_if("data/processed/consultations/consultation_audit_source_coverage.csv")
by_year_src     <- read_if("data/processed/consultations/consultation_audit_2023_2025_by_year_source.csv")
qual_highlights      <- read_if("data/processed/consultations/qualitative_highlights.csv")
impact_case_studies  <- read_if("data/processed/consultations/impact_case_studies.csv")

kpi <- function(key) {
  if (nrow(audit_summary) == 0) return("—")
  x <- audit_summary |> filter(metric_id == key) |> pull(value)
  if (length(x) == 0) "—" else comma(x[[1]])
}

no_data_msg <- function() {
  cat("*Data file not loaded. Run the ETL scripts to populate this chart.*\n")
}

# ── Pre-computed alt text for charts ─────────────────────────
if (data_loaded) {
  .yt <- rec_year_team |> mutate(group = recode(group, Datasquad = "walk-in sign-ins", DSC = "scheduled appointments", Other = "other"))
  alt_year <- sprintf(
    "Horizontal stacked bar chart: consultation records file by year and record type, %s. %s. 2024 covers January to early June only.",
    rec_window_label,
    paste(sprintf("%d: %s", .yt$year, paste(.yt$group, .yt$records)), collapse = "; ")
  )
  alt_group <- sprintf(
    "Bar chart: consultation records by team, %s. %s.",
    rec_window_label, paste(rec_team$group, rec_team$records, sep = " = ", collapse = "; ")
  )
} else {
  alt_year  <- "Consultation records by year (data not loaded)"
  alt_group <- "Consultation records by team (data not loaded)"
}

data_range_label <- if (nrow(audit_coverage) > 0) {
  yr_min <- format(min(as.Date(audit_coverage$min_date)), "%Y")
  yr_max <- format(max(as.Date(audit_coverage$max_date)), "%Y")
  if (yr_min == yr_max) yr_min else paste0(yr_min, "\u2013", yr_max)
} else {
  "All Time"
}

# YoY comparison: 2023 → 2024 (two complete calendar years in this audit)
# 2025 data is partial (libinsights export ends March 2025) — excluded from YoY
yoy_fmt <- function(n_recent, n_prior, yr_recent = 2024, yr_prior = 2023) {
  if (!length(n_recent) || !length(n_prior) || is.na(n_recent) || is.na(n_prior) || n_prior == 0) return("")
  pct   <- round((n_recent - n_prior) / n_prior * 100)
  arrow <- if (pct > 0) "\u2191" else if (pct < 0) "\u2193" else "\u2192"
  sprintf("%s %+d%% \u00b7 %s in %d vs %s in %d",
          arrow, pct, comma(n_recent), yr_recent, comma(n_prior), yr_prior)
}

yoy_direct <- local({
  if (nrow(by_year_src) == 0) return("")
  d   <- by_year_src |> filter(source != "trello") |>
         group_by(year) |> summarise(n = sum(n), .groups = "drop")
  yoy_fmt(d$n[d$year == 2024], d$n[d$year == 2023])
})

yoy_tasks <- local({
  if (nrow(audit_task_mode) == 0) return("")
  d <- audit_task_mode |> filter(task_metric == "trello_reportable_task_cards")
  yoy_fmt(d$n[d$year == 2024], d$n[d$year == 2023])
})
```

*For DSC staff and library administration — direct consultation counts, project task metrics, and patron breakdowns.*

## Recorded Consultations Since 2017

```{r series_setup, include=FALSE}
consult_series <- read_if("data/processed/canonical/consultations_by_year.csv")
series_total <- read_if("data/processed/canonical/headline_aggregates.csv") |>
  filter(metric_id == "consultations_recorded_series") |> pull(display)
if (length(series_total) == 0) series_total <- "—"
```

DSC recorded **`r series_total` research consultations** from 2017 to September 2026 (claim-register #10). The series counts scheduled appointments, with DSC staff and DataSquad consultants alike; DataSquad's in-person walk-in sign-ins are not included. Records are incomplete after mid-2024: Calendly ended, LibCal wasn't in use until May 2025, and two reorganizations and staff departures cut logging. Because of that gap, this series cannot measure how demand changed after 2023; LibCal shows the work continued. It is a different series from the 2023-2025 audit below, which adds DataSquad walk-ins and uses its own deduplication.

```{r series_table, echo=FALSE}
if (nrow(consult_series) > 0) {
  consult_series |>
    transmute(Year = as.character(year), `Recorded consultations` = display,
              Note = if_else(is_floor, "Floor (recording lapsed)", "")) |>
    knitr::kable(align = c("l", "r", "l"))
}
```

## 2023-2025 Consultation Audit (a separate series)

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*

::: {.kpi-row .grid}

::: {.g-col-6 .g-col-md-3 .kpi-card}
[Direct Consultations]{.kpi-label}
[`r kpi("direct_consults_combined_deduped_count")`]{.kpi-value}
[DSC + DataSquad combined; canceled excluded. Recording coverage differs by year, so year-to-year change is not shown]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[DSC Staff]{.kpi-label}
[`r kpi("dsc_direct_consults_libinsights_count")`]{.kpi-value}
[LibInsight scheduled exports]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[DataSquad]{.kpi-label}
[`r kpi("datasquad_direct_consults_count")`]{.kpi-value}
[Walk-in sign-ins]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[Task Cards (Reportable)]{.kpi-label}
[`r kpi("datasquad_trello_task_cards_reportable_count")`]{.kpi-value}
[Trello; backlog/admin lists excluded. Trello tracking began in 2023]{.kpi-note}
:::

:::

## Direct Consultations

::: {.panel-tabset}

### By Year (consultation records file, `r rec_window_label`)

This chart counts rows in the consultation records file that also feeds the department and affiliation breakdowns. The file runs from `r rec_window_label`, so 2024 covers only part of the year. Unlike the series at the top of this page, which counts scheduled appointments (with DSC staff and DataSquad consultants alike), this file also includes DataSquad's in-person walk-in sign-ins. That is why its 2024 total is larger: for January to May 2024 the scheduled appointments in both sources agree (166 here, 165 in the series).

```{r chart_year, echo=FALSE, fig.alt=alt_year, fig.height=3.6}
if (data_loaded && nrow(rec_year_team) > 0) {
  type_levels <- c("Scheduled appointments", "Other", "DataSquad walk-in sign-ins")
  rec_year_team |>
    mutate(type = factor(recode(group, DSC = "Scheduled appointments", Datasquad = "DataSquad walk-in sign-ins"),
                         levels = rev(type_levels)),
           year = factor(year, levels = rev(sort(unique(year))))) |>
    ggplot(aes(y = year, x = records, fill = type)) +
    geom_col(width = 0.65, color = "white", linewidth = 0.4) +
    scale_fill_manual(values = c("Scheduled appointments" = "#2774AE", "Other" = "#8C8279",
                                 "DataSquad walk-in sign-ins" = "#FFB81C"),
                      breaks = type_levels, name = NULL) +
    labs(y = NULL, x = "Records",
         title = "Consultation records file by record type (2024: January to early June only)") +
    theme_minimal() +
    theme(panel.grid.major.y = element_blank(), legend.position = "bottom")
} else { no_data_msg() }
```

### By Department

```{r chart_dept, echo=FALSE}
if (data_loaded) {
  dept_counts <- rec_dept |>
    rename(Department = department, Consultations = records)

  sd_dept <- SharedData$new(dept_counts)

  # Filter widget + linked chart + linked table
  filter_select("dept_filter", "Search department:", sd_dept, ~Department,
                multiple = FALSE)

  bscols(
    plot_ly(sd_dept, x = ~Consultations, y = ~reorder(Department, Consultations),
            type = "bar", orientation = "h",
            marker = list(color = "#005587"),
            hovertemplate = "<b>%{y}</b><br>Consultations: %{x}<extra></extra>") |>
      layout(yaxis = list(title = ""),
             xaxis = list(title = "Consultations"),
             margin = list(l = 5)),
    DT::datatable(sd_dept, rownames = FALSE,
                  colnames = c("Department", "Consultations"),
                  options = list(pageLength = 10, dom = "tp"),
                  caption = paste0("Departments, consultation records file, ", rec_window_label, ".", suppressed_note(rec_window$suppressed_departments, "departments")))
  )
} else { no_data_msg() }
```

### By Affiliation

```{r chart_affil, echo=FALSE}
if (data_loaded) {
  affil_counts <- rec_affil |>
    rename(Affiliation = ucla_affiliation, Consultations = records)

  sd_affil <- SharedData$new(affil_counts)

  filter_checkbox("affil_filter", "Filter affiliation:", sd_affil, ~Affiliation)

  bscols(
    plot_ly(sd_affil, x = ~Consultations, y = ~reorder(Affiliation, Consultations),
            type = "bar", orientation = "h",
            marker = list(color = "#005587"),
            hovertemplate = "<b>%{y}</b><br>Consultations: %{x}<extra></extra>") |>
      layout(yaxis = list(title = ""),
             xaxis = list(title = "Consultations"),
             margin = list(l = 5)),
    DT::datatable(sd_affil, rownames = FALSE,
                  colnames = c("Affiliation", "Consultations"),
                  options = list(pageLength = 10, dom = "t"),
                  caption = paste0("Affiliations, consultation records file, ", rec_window_label, ".", suppressed_note(rec_window$suppressed_affiliations, "affiliations")))
  )
} else { no_data_msg() }
```

### By Team

```{r chart_group, echo=FALSE, fig.alt=alt_group}
if (data_loaded) {
  rec_team |>
    rename(n = records) |>
    ggplot(aes(x = group, y = n, fill = group)) +
    geom_col() +
    scale_fill_manual(values = c("DSC" = "#003B5C", "Datasquad" = "#FFB81C")) +
    labs(x = NULL, y = "Consultations") +
    theme_minimal() +
    theme(legend.position = "none")
} else { no_data_msg() }
```

:::

## Project Work & Task Metrics

These metrics represent follow-on project work assigned to DataSquad members. They are **not** direct consultations.

::: {.panel-tabset}

### Summary

```{r task_table, echo=FALSE}
if (nrow(audit_summary) == 0) {
  cat("*Run `Rscript src/etl/consultation_audit_summary.R` to generate task metrics.*\n")
} else {
  audit_summary |>
    filter(metric_id %in% c(
      "direct_consults_combined_deduped_count",
      "datasquad_trello_task_cards_reportable_count",
      "datasquad_trello_task_activity_comment_weighted_reportable_count",
      "datasquad_total_with_trello_reportable_count"
    )) |>
    transmute(
      Metric = recode(metric_id,
        direct_consults_combined_deduped_count =
          "Direct Consultations (Combined, Deduped)",
        datasquad_trello_task_cards_reportable_count =
          "DataSquad Task Cards (Reportable)",
        datasquad_trello_task_activity_comment_weighted_reportable_count =
          "DataSquad Task Activity (Reportable, Comment-Weighted)",
        datasquad_total_with_trello_reportable_count =
          "DataSquad Direct + Reportable Task Activity"
      ),
      Value = comma(value)
    ) |>
    knitr::kable(align = c("l", "r"))
}
```

### Task Activity by Year

```{r task_year, echo=FALSE, fig.alt="Two bar charts of DataSquad project work logged in Trello by year, 2023 to 2025, reportable tasks only: task cards and comment-weighted activity. Values are labeled on each bar. Trello tracking began in 2023, so early years undercount."}
if (nrow(audit_task_mode) == 0) {
  cat("*Task-mode audit file not found.*\n")
} else {
  audit_task_mode |>
    filter(task_metric %in% c("trello_reportable_task_cards", "trello_reportable_comment_weighted")) |>
    mutate(metric_family = if_else(grepl("cards", task_metric),
                                   "Reportable task cards", "Comment-weighted activity")) |>
    ggplot(aes(x = factor(year), y = n)) +
    geom_col(fill = "#2774AE", width = 0.6) +
    geom_text(aes(label = n), vjust = -0.4, size = 3.5) +
    facet_wrap(~metric_family, scales = "free_y") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
    labs(
      x = NULL, y = NULL,
      title = "DataSquad project work logged in Trello (tracking began in 2023)",
      caption = "Reportable tasks only (excludes backlog, admin, and archived lists). Growth from 2023 partly reflects adoption of Trello tracking."
    ) +
    theme_minimal() +
    theme(panel.grid.major.x = element_blank(), panel.grid.minor = element_blank())
}
```

:::

## Research Impact

These are illustrative cases selected to show different forms of contribution, not a representative sample of DSC and DataSquad's work. They show how support extends into sustained technical collaboration beyond single consultations.

```{r impact, echo=FALSE, results='asis'}
if (nrow(impact_case_studies) == 0) {
  cat("*Add rows to `data/processed/consultations/impact_case_studies.csv` to render case studies here.*\n")
} else {
  impact_case_studies |>
    filter(coalesce(include_in_dashboard, TRUE)) |>
    arrange(display_order) |>
    pwalk(function(case_title, partners, summary,
                   measurable_outcomes, impact_signals, ...) {
      cat("### ", case_title, "\n\n", sep = "")
      if (!is.na(partners) && partners != "")
        cat("**Partners:** ", partners, "\n\n", sep = "")
      if (!is.na(summary) && summary != "")
        cat(summary, "\n\n")
      if (!is.na(measurable_outcomes) && measurable_outcomes != "")
        cat("**Measurable outcomes:** ", measurable_outcomes, "\n\n", sep = "")
      if (!is.na(impact_signals) && impact_signals != "")
        cat("**Impact signals:** ", impact_signals, "\n\n", sep = "")
    })
}
```

```{r qualitative, echo=FALSE}
if (nrow(qual_highlights) > 0) {
  display_cols <- c("period_label", "workstream", "audience",
                    "request_theme", "work_performed", "outcome")
  qual_highlights |>
    select(all_of(intersect(display_cols, names(qual_highlights)))) |>
    rename_with(~str_replace_all(., "_", " ")) |>
    knitr::kable(align = "l")
}
```

---

**See also:** [Instruction <span aria-hidden="true">→</span>](instruction.qmd) workshop attendance · [Infrastructure <span aria-hidden="true">→</span>](infrastructure.qmd) Dataverse and AWS metrics · [About & methodology](about.qmd)

---

::: {.callout-note collapse="true" .data-notes title="Data notes"}
- **Direct consultations**: LibInsight scheduled exports + manual logs (DSC staff); DataSquad walk-in sign-in forms
- **Canceled appointments** excluded from all direct consultation counts
- **Deduplication**: manual LibInsight logs matched against scheduled exports using a same-provider/same-minute heuristic; ~`r kpi("libinsight_manual_logs_heuristic_duplicate_count")` duplicates removed
- **Task metrics** (Trello cards, comment-weighted activity) are shown separately — not comparable to consultation counts
- **Reportable filter**: excludes `Inbox`, `Future Projects/Ideas`, `Team Info` lists; includes archived completed cards
- **YoY comparisons** (2023 → 2024): direct consultation decline reflects shifting data sources and intake channels, not necessarily service volume; task card growth reflects expanded Trello adoption in 2024, not a proportional workload increase
- Full rules: [Metrics Governance](metrics-governance.md) · [About & methodology](about.qmd)
:::

```{r coverage, echo=FALSE}
if (nrow(audit_coverage) > 0) {
  audit_coverage |>
    mutate(min_date = as.Date(min_date), max_date = as.Date(max_date)) |>
    arrange(source, subtype) |>
    transmute(
      Source        = source,
      `Record type` = subtype,
      Rows          = comma(rows),
      `Date range`  = paste(format(min_date, "%Y-%m-%d"), "\u2013",
                            format(max_date, "%Y-%m-%d"))
    ) |>
    DT::datatable(
      rownames = FALSE,
      options  = list(pageLength = 10, dom = "tp"),
      caption  = "Source coverage: rows and date ranges per data system"
    )
}
```

::: {.callout-note collapse="true" title="Cite this page"}
UCLA Library Data Science Center. (`r year(Sys.Date())`). *Consulting Statistics*. Retrieved from <https://ucla-data-science-center.github.io/dsc-stats-reports/consulting.html>
:::

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
