---
title: "DSC Statistics Dashboard"
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)

audit_summary <- local({
  p <- "data/processed/consultations/consultation_audit_2023_2025_summary_tagged.csv"
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else tibble()
})

audit_coverage <- local({
  p <- "data/processed/consultations/consultation_audit_source_coverage.csv"
  if (file.exists(p)) read_csv(p, show_col_types = FALSE) else tibble()
})

kpi <- function(key) {
  if (nrow(audit_summary) == 0) return("—")
  x <- audit_summary |> filter(metric_id == key) |> pull(value)
  if (length(x) == 0) "—" else comma(x[[1]])
}

direct_total     <- kpi("direct_consults_combined_deduped_count")
dsc_direct       <- kpi("dsc_direct_consults_libinsights_count")
datasquad_direct <- kpi("datasquad_direct_consults_count")
task_cards       <- kpi("datasquad_trello_task_cards_reportable_count")

data_range_label <- if (nrow(audit_coverage) > 0) {
  yr_min <- format(min(as.Date(audit_coverage$min_date)), "%Y")
  yr_max <- format(max(as.Date(audit_coverage$max_date)), "%Y")
  if (yr_min == yr_max) yr_min else paste0(yr_min, "\u2013", yr_max)
} else {
  "All Time"
}
```

UCLA Library Data Science Center unified service metrics — consultations, instruction, and infrastructure in one place.

## At a Glance (`r data_range_label`)

::: {.kpi-row .grid}

::: {.g-col-6 .g-col-md-3 .kpi-card}
[Direct Consultations]{.kpi-label}
[`r direct_total`]{.kpi-value}
[DSC + DataSquad combined; canceled excluded]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[DSC Staff]{.kpi-label}
[`r dsc_direct`]{.kpi-value}
[LibInsight scheduled exports]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[DataSquad]{.kpi-label}
[`r datasquad_direct`]{.kpi-value}
[Walk-in sign-ins]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-3 .kpi-card}
[Task Cards (Reportable)]{.kpi-label}
[`r task_cards`]{.kpi-value}
[Trello; backlog/admin lists excluded]{.kpi-note}
:::

:::

## Dashboard Sections

::: {.grid}

::: {.g-col-4 .card .p-3}
### [Consulting](consulting.qmd)
One-on-one research support — DSC staff appointments and DataSquad walk-ins.
:::

::: {.g-col-4 .card .p-3}
### [Instruction](instruction.qmd)
Workshops, Carpentries events, and community-led training sessions.
:::

::: {.g-col-4 .card .p-3}
### [Infrastructure](infrastructure.qmd)
AWS cloud costs and UCLA Dataverse repository growth and download metrics.
:::

:::

## Reading These Metrics

- **Direct consultations** = one interaction record (appointment or walk-in). Canceled appointments excluded.
- **Task cards** = Trello cards assigned to DataSquad project work. Not equivalent to consultations.
- **Task activity (comment-weighted)** = comment count on task cards — a workload proxy, not hours or outputs.

::: {.callout-note collapse="true" title="Data notes"}
- Sources: LibInsight (scheduled exports + manual logs), DataSquad walk-in sign-in forms, Trello (historical task tracking)
- Canceled appointments excluded from all direct consultation counts
- Manual LibInsight logs deduplicated against scheduled exports using a same-provider/same-minute heuristic
- Task metrics (Trello cards, comment-weighted activity) are **not** consultations — shown separately throughout
- Full counting rules: [Metrics Governance](metrics-governance.md) · [About & methodology](about.qmd)
:::

```{r coverage, echo=FALSE, results='asis'}
if (nrow(audit_coverage) > 0) {
  snap <- audit_coverage |>
    mutate(max_date = as.Date(max_date)) |>
    arrange(desc(max_date)) |>
    slice_head(n = 4)
  cat(
    "**Data coverage:** ",
    paste(
      sprintf("%s through %s", snap$source, format(snap$max_date, "%Y-%m-%d")),
      collapse = "; "
    ),
    "\n"
  )
}
```

---

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
