---
title: "Instruction Statistics"
format:
  html:
    toc: true
    code-fold: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
source("src/etl/instruction_aggregate.R")
source("src/etl/suppression.R")

# Reconciled headline figures (dsc-stats-integration, see claim-register.md #9, #12)
headline <- if (file.exists("data/processed/canonical/headline_aggregates.csv")) {
  read_csv("data/processed/canonical/headline_aggregates.csv", show_col_types = FALSE)
} else tibble(metric_id = character(), display = character())
hl <- function(key) {
  x <- headline$display[headline$metric_id == key]
  if (length(x) == 0) "—" else x[[1]]
}

# Registered seats by program and year (claim-register.md #12)
seats <- if (file.exists("data/processed/canonical/instruction_seats_by_program_year.csv")) {
  read_csv("data/processed/canonical/instruction_seats_by_program_year.csv", show_col_types = FALSE)
} else tibble()

# ── Workshop data (full attendee source; graceful fallback) ───
data_loaded    <- FALSE
ucla_workshops <- tibble()

raw_workshop_path <- "data/raw/instruction/dsc_workshops.csv"
if (file.exists(raw_workshop_path)) {
  ucla_workshops <- load_instruction_attendee_data(raw_workshop_path)
  data_loaded <- TRUE
} else {
  for (p in c(
    "data/processed/instruction/ucla_workshops.rda",
    "data/instruction/ucla_workshops.rda",
    "data/consultations/Data/ucla_workshops.rda"
  )) {
    if (file.exists(p)) {
      load(p)
      data_loaded <- TRUE
      break
    }
  }
}

if (!data_loaded) {
  message("ucla_workshops.rda not found — charts will show placeholder text.")
}

instruction_attendance <- if (data_loaded) {
  load_instruction_attendance(ucla_workshops)
} else {
  tibble()
}

no_data_msg <- function() {
  cat("*Data file not loaded. Run the ETL scripts to populate this chart.*\n")
}

# ── KPIs ──────────────────────────────────────────────────────
fmt <- function(x) if (is.na(x) || length(x) == 0) "—" else comma(x)

kpi_total <- if (data_loaded) sum(instruction_attendance$attendance_count) else NA_integer_
kpi_ucla  <- if (data_loaded) {
  sum(instruction_attendance$attendance_count[instruction_attendance$institution == "UCLA"], na.rm = TRUE)
} else NA_integer_
kpi_ucla_held <- if (data_loaded && "ucla_held" %in% names(instruction_attendance)) {
  sum(instruction_attendance$attendance_count[instruction_attendance$ucla_held], na.rm = TRUE)
} else NA_integer_
kpi_unknown_affiliation <- if (data_loaded) {
  sum(instruction_attendance$attendance_count[is.na(instruction_attendance$institution)])
} else NA_integer_
kpi_organizational_parent <- if (data_loaded) {
  sum(instruction_attendance$attendance_count[
    instruction_attendance$institution == "UCLA" &
      !is.na(instruction_attendance$organizational_parent)
  ], na.rm = TRUE)
} else NA_integer_
kpi_years <- if (data_loaded) n_distinct(year(instruction_attendance$date), na.rm = TRUE) else NA_integer_

# ── Date range label and YoY comparison ───────────────────────
data_range_label <- if (data_loaded) {
  yr_min <- min(year(instruction_attendance$date), na.rm = TRUE)
  yr_max <- max(year(instruction_attendance$date), na.rm = TRUE)
  if (yr_min == yr_max) as.character(yr_min) else paste0(yr_min, "\u2013", yr_max)
} else "All Time"

yoy_attendance <- if (data_loaded) {
  by_year <- instruction_attendance |>
    filter(!is.na(date)) |>
    mutate(year = year(date)) |>
    summarise(n = sum(attendance_count), .by = year) |>
    arrange(desc(year)) |>
    head(2)
  if (nrow(by_year) >= 2) {
    n_recent <- by_year$n[1];  yr_recent <- by_year$year[1]
    n_prior  <- by_year$n[2];  yr_prior  <- by_year$year[2]
    pct   <- round((n_recent - n_prior) / n_prior * 100)
    arrow <- if (pct > 0) "\u2191" else if (pct < 0) "\u2193" else "\u2192"
    sprintf("%s %+d%% \u00b7 %s in %d vs %s in %d",
            arrow, pct, comma(n_recent), yr_recent, comma(n_prior), yr_prior)
  } else ""
} else ""

# ── Pre-computed alt text ─────────────────────────────────────
if (data_loaded) {
  .yr <- instruction_attendance |> filter(!is.na(date)) |>
    mutate(Year = year(date)) |>
    summarise(n = sum(attendance_count), .by = Year) |> arrange(Year)
  alt_year <- sprintf(
    "Horizontal bar chart: workshop attendance by year from %d to %d. %s.",
    min(.yr$Year), max(.yr$Year),
    paste(.yr$Year, .yr$n, sep = ": ", collapse = "; ")
  )
  .dept <- instruction_attendance |>
    filter(!is.na(standardized_department), institution == "UCLA") |>
    summarise(n = sum(attendance_count), .by = standardized_department) |>
    arrange(desc(n)) |> head(15)
  alt_dept <- if (nrow(.dept) > 0) sprintf(
      "Horizontal bar chart: top %d UCLA departments by workshop attendance. Highest: %s (%d).",
      nrow(.dept), .dept$standardized_department[1], .dept$n[1]
    ) else "Workshop attendance by UCLA department (no known departments)"
  .affil <- instruction_attendance |> filter(!is.na(status)) |>
    summarise(n = sum(attendance_count), .by = status) |>
    arrange(desc(n)) |> head(5)
  alt_affil <- if (nrow(.affil) > 0) sprintf(
      "Horizontal bar chart: workshop attendance by participant affiliation. Top group: %s (%d).",
      .affil$status[1], .affil$n[1]
    ) else "Workshop attendance by affiliation (no known affiliations)"
} else {
  alt_year  <- "Workshop attendance by year (data not loaded)"
  alt_dept  <- "Top departments by workshop attendance (data not loaded)"
  alt_affil <- "Attendance by affiliation (data not loaded)"
}
```

*Evidence page: registered seats and workshop attendance from attendee-level and validated aggregate sources, with their definitions and limits.*

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*

::: {.callout-note}
## Headline figures since 2017

DSC instruction is counted in registered seats: one registration for one session, not a unique person and not an attendance. The series runs from May 2017 through September 2026 (2026 is partial) and includes the UC-wide series DSC co-founded and co-delivers, so UCLA participants are a subset. DSC offered `r hl("instruction_sessions_lifetime")` workshop sessions from May 2017 through May 2026. The sections after the seat chart cover DSC's attendee-level file only.
:::

## Registered Seats by Program, 2017-2026

```{r seats_setup, include=FALSE}
source("src/etl/seats_chart.R")
alt_seats <- seats_alt_text(seats)
```

::: {.grid}

::: {.g-col-12 .g-col-md-8}

```{r seats_chart, echo=FALSE, fig.alt=alt_seats, fig.width=7, fig.height=4.2, fig.cap=seats_caption}
if (nrow(seats) > 0) seats_chart(seats)
```

:::

::: {.g-col-12 .g-col-md-4 .seat-total}

[`r hl("instruction_registered_seats")`]{.kpi-value}

[Registered seats, May 2017 to September 2026 (2026 partial)]{.kpi-label}

[Registrations, not unique people or attendance.]{.kpi-note}

:::

:::

```{r seats_table, echo=FALSE}
if (nrow(seats) > 0) {
  seat_wide <- seats |>
    select(Program = program, year, registered_seats) |>
    arrange(year) |>
    pivot_wider(names_from = year, values_from = registered_seats) |>
    mutate(Total = rowSums(across(-Program), na.rm = TRUE))
  seat_total <- seat_wide |>
    summarise(across(-Program, \(x) sum(x, na.rm = TRUE))) |>
    mutate(Program = "Total", .before = 1)
  bind_rows(seat_wide, seat_total) |>
    mutate(across(-Program, \(x) if_else(is.na(x), "", comma(x)))) |>
    knitr::kable(
      align = c("l", rep("r", ncol(seat_wide) - 1)),
      caption = "Registered seats (one registration for one session), May 2017 to September 2026. 2026 is partial."
    )
}
```

Attended seats are shown only where attendance was recorded for the same series and dates (Zoom at 30 minutes or more, or an attendance sheet). Coverage differs by series, so each ratio of attended to registered applies to its own row only, not to other series or to the total.

```{r seats_attended, echo=FALSE}
if (nrow(seats) > 0) {
  seats |>
    filter(!is.na(attended_seats)) |>
    transmute(Program = program, Year = as.character(year),
              `Registered seats` = comma(registered_seats),
              `Attended seats` = comma(attended_seats),
              Note = coalesce(attended_note, "")) |>
    knitr::kable(align = c("l", "l", "r", "r", "l"),
                 caption = "Registered and attended seats where attendance was recorded")
}
```

::: {.callout-note collapse="true" title="How the seat series is counted"}
- **Unit**: registered seats, one per person per session signed up. One unit is used for each series: UC Carpentries and Library Carpentry count registered seats, Love Data Week counts registered seats (best effort), GIS Week counts registrations, and DSC-run workshops use the larger of a rebuild from named registration sources and the published aggregate, per year
- **Window**: May 2017 through September 2026. 2026 is partial because the Fall 2026 UC Carpentries series is in progress
- **Why it differs from the earlier 17,466 (May 2017 to April 2024)**: earlier Carpentries figures counted attended seats or people, not registered seats; the window now runs through 2026; and Love Data Week 2021 is counted on the same UC-wide basis as 2022-2024 (1,529 registered seats, where the earlier series counted 177 from DSC-hosted sessions only)
- **Open items**: Library Carpentry May 2026 attended is 428 here (rejoins merged by name) against 453 in the earlier series; UC Carpentries September 2024 attended is 1,036 Zoom rows at 30 minutes or more, 1,027 after merging rejoins. Named sources for general DSC workshops in 2017-2019 are incomplete, so those years use the published aggregate where it is larger
- **Aggregates only**: row-level registration and attendance data stay outside this repository
:::

## Attendance records available for analysis (`r data_range_label`)

::: {.kpi-row .grid}

::: {.g-col-6 .g-col-md-4 .kpi-card}
[Total Attendance Records]{.kpi-label}
[`r fmt(kpi_total)`]{.kpi-value}
[Attendee-level file only (May 2017 to Sept 2024; 2024 partial); see headline figures above]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[UCLA-Affiliated]{.kpi-label}
[`r fmt(kpi_ucla)`]{.kpi-value}
[institution == "UCLA"]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[UCLA-Held]{.kpi-label}
[`r fmt(kpi_ucla_held)`]{.kpi-value}
[Attendee-events, May 2017–Apr 2024: UCLA-affiliated, plus unknown-affiliation records from 2017–2020 sessions held at UCLA]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[Years with Data]{.kpi-label}
[`r fmt(kpi_years)`]{.kpi-value}
[Distinct calendar years]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[UCLA Affiliation Unknown]{.kpi-label}
[`r fmt(kpi_unknown_affiliation)`]{.kpi-value}
[Included in totals; excluded from affiliation, department, and school metrics]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[School, division, or other unit identified]{.kpi-label}
[`r fmt(kpi_organizational_parent)`]{.kpi-value}
[`r percent(kpi_organizational_parent / kpi_ucla, accuracy = 0.1)` of validated UCLA attendance]{.kpi-note}
:::

:::

## Workshop Attendance

::: {.panel-tabset}

### By Year (attendee-level records)

Attendee-level records only, May 2017 to September 2024 (2024 is partial). This is a different series from the registered-seat chart above.

```{r chart_year, echo=FALSE, fig.alt=alt_year}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(date)) |>
    mutate(Year = year(date)) |>
    summarise(n = sum(attendance_count), .by = Year) |>
    ggplot(aes(x = factor(Year), y = n)) +
    geom_col(fill = "#005587") +
    coord_flip() +
    labs(x = NULL, y = "Attendance") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

### UC-Wide Series

```{r uc_series_table, echo=FALSE}
if (data_loaded) {
  instruction_attendance |>
    mutate(
      Series = case_when(
        str_detect(str_to_lower(coalesce(event, "")), "love data|ldw") ~ "Love Data Week",
        str_detect(str_to_lower(coalesce(event, "")), "gis week|uc gis") ~ "GIS Week",
        TRUE ~ NA_character_
      ),
      Year = year(date)
    ) |>
    filter(!is.na(Series), !is.na(Year)) |>
    summarise(Attendance = sum(attendance_count), .by = c(Series, Year)) |>
    arrange(Series, Year) |>
    knitr::kable(
      align = c("l", "r", "r"),
      caption = "Attendance instances for the UC-wide series in the attendee-level file. The reconciled registered-seat counts, including Love Data Week 2021 at 1,529, are in the seat table above."
    )
}
```

### By Department

```{r chart_dept, echo=FALSE, fig.alt=alt_dept}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(standardized_department), institution == "UCLA") |>
    summarise(n = sum(attendance_count), .by = standardized_department) |>
    filter(n >= 5) |>
    arrange(desc(n)) |>
    head(15) |>
    ggplot(aes(x = reorder(standardized_department, n), y = n)) +
    geom_col(fill = "#005587") +
    coord_flip() +
    labs(x = NULL, y = "Attendance", title = "Top 15 Departments (UCLA attendees)") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

```{r dept_table, echo=FALSE}
if (data_loaded) {
  dept_tbl <- instruction_attendance |>
    filter(!is.na(standardized_department), institution == "UCLA") |>
    summarise(n = sum(attendance_count), .by = standardized_department) |>
    arrange(desc(n)) |>
    suppress_small_cells("standardized_department", "n")
  dept_tbl |>
    rename(Department = standardized_department, Attendance = n) |>
    DT::datatable(
      rownames = FALSE,
      options  = list(pageLength = 10, dom = "ftp"),
      caption  = paste0("UCLA departments. Search or sort to find a specific unit.",
                        suppression_note(dept_tbl, "departments"))
    )
}
```

### By Organizational Parent

```{r organizational_parent_table, echo=FALSE}
if (data_loaded) {
  instruction_attendance |>
    filter(institution == "UCLA", !is.na(organizational_parent)) |>
    summarise(Attendance = sum(attendance_count), .by = c(organizational_parent, parent_type)) |>
    arrange(desc(Attendance)) |>
    suppress_small_cells("organizational_parent", "Attendance") -> parent_tbl
  parent_tbl |>
    rename(`Organizational parent` = organizational_parent, `Parent type` = parent_type) |>
    DT::datatable(
      rownames = FALSE,
      options = list(pageLength = 15, dom = "ftp"),
      caption = paste0("Validated UCLA attendance by broader organizational parent.",
                       suppression_note(parent_tbl, "parents"))
    )
}
```

### By Affiliation

```{r chart_affil, echo=FALSE, fig.alt=alt_affil}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(status)) |>
    summarise(n = sum(attendance_count), .by = status) |>
    filter(n >= 5) |>
    arrange(desc(n)) |>
    head(5) |>
    ggplot(aes(x = reorder(status, n), y = n)) +
    geom_col(fill = "#005587") +
    coord_flip() +
    labs(x = NULL, y = "Attendance") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

```{r affil_table, echo=FALSE}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(status)) |>
    summarise(n = sum(attendance_count), .by = status) |>
    arrange(desc(n)) |>
    suppress_small_cells("status", "n") -> affil_tbl
  affil_tbl |>
    rename(Affiliation = status, Attendance = n) |>
    DT::datatable(
      rownames = FALSE,
      options  = list(pageLength = 10, dom = "ftp"),
      caption  = paste0("Participant affiliations.", suppression_note(affil_tbl, "affiliations"))
    )
}
```

:::

**See also:** [Consulting <span aria-hidden="true">→</span>](consulting.qmd) one-on-one consultations · [Infrastructure <span aria-hidden="true">→</span>](infrastructure.qmd) Dataverse and AWS metrics · [About & methodology](about.qmd)

---

[]{#data-notes}

::: {.callout-note collapse="true" .data-notes title="Data notes"}
- **Sources**: Consolidated workshop attendee records (including UC-wide GIS Week and Love Data Week) and privacy-safe aggregate workshop summaries
- **Unit**: Attendance instances. LibInsight uses one attendee-event row; aggregate sources provide a validated count for each event. A person attending three workshops contributes three attendance instances
- **UCLA attendees**: filtered on `institution == "UCLA"`. Records with unknown institution remain in overall/year/topic totals but are excluded from UCLA, department, and school metrics
- **UCLA-held**: UCLA-affiliated records plus records with unknown institution from 2017–2020 sessions classified as held at UCLA (`data/reference/instruction/session_venue_2017_2020.tsv`, reviewed by the DSC director, 2026-09). Institution was not collected before 2020, so the UCLA-affiliated figure reads zero for 2017–2019 although that teaching was on campus. A UCLA-held session can include a few outside attendees. Sessions classified as mixed (co-hosted or UC-wide) or held elsewhere are excluded. This metric does not feed department or school counts
- **Department**: `standardized_department` field, mapped from raw free-text entries; unmapped values appear as-is
- **Organizational parent**: versioned mapping from validated UCLA departments to College divisions, professional schools, the Library, UCLA Health, institutes, or administrative groups. Missing and unresolved departments remain unknown
- **School/division**: not inferred when the source lacks a validated department. Unknown is never treated as non-UCLA
- **UC-wide series**: GIS Week and Love Data Week are included from the consolidated attendee source; their upstream workbooks are retained as provenance and are not added again
- **Registered seats (headline)**: a separate reconciled series from `data/processed/canonical/instruction_seats_by_program_year.csv`, counted as described under "How the seat series is counted". It is not the sum of the attendee-level tables on this page
- **Small cells**: department, organizational-parent, and affiliation breakdowns fold any unit with fewer than 5 into one combined row, which is omitted if it is also under 5
- Full rules: [Metrics Governance](metrics-governance.md) · [About & methodology](about.qmd)
:::

```{r coverage, echo=FALSE, results='asis'}
if (data_loaded && !is.null(instruction_attendance$date)) {
  instruction_attendance |>
    mutate(`Record type` = if_else(record_granularity == "aggregate", "event_summary", "attendee_event")) |>
    summarise(
      Rows = n(),
      Attendance = sum(attendance_count),
      `Date range` = paste(format(min(date, na.rm = TRUE), "%Y-%m-%d"), "\u2013",
                           format(max(date, na.rm = TRUE), "%Y-%m-%d")),
      .by = c(source_system, `Record type`)
    ) |>
    rename(Source = source_system) |>
    mutate(Rows = scales::comma(Rows), Attendance = scales::comma(Attendance)) |>
    knitr::kable(align = c("l", "l", "r", "r", "l"))
}
```

::: {.callout-note collapse="true" title="Cite this page"}
UCLA Library Data Science Center. (`r year(Sys.Date())`). *Instruction Statistics*. Retrieved from <https://ucla-data-science-center.github.io/dsc-stats-reports/instruction.html>
:::

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
