---
title: "Instruction Statistics"
format:
  html:
    toc: true
    code-fold: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
source("src/etl/instruction_aggregate.R")

# ── Workshop data (full attendee source; graceful fallback) ───
data_loaded    <- FALSE
ucla_workshops <- tibble()

raw_workshop_path <- "data/raw/instruction/dsc_workshops.csv"
if (file.exists(raw_workshop_path)) {
  ucla_workshops <- load_instruction_attendee_data(raw_workshop_path)
  data_loaded <- TRUE
} else {
  for (p in c(
    "data/processed/instruction/ucla_workshops.rda",
    "data/instruction/ucla_workshops.rda",
    "data/consultations/Data/ucla_workshops.rda"
  )) {
    if (file.exists(p)) {
      load(p)
      data_loaded <- TRUE
      break
    }
  }
}

if (!data_loaded) {
  message("ucla_workshops.rda not found — charts will show placeholder text.")
}

instruction_attendance <- if (data_loaded) {
  load_instruction_attendance(ucla_workshops)
} else {
  tibble()
}

no_data_msg <- function() {
  cat("*Data file not loaded. Run the ETL scripts to populate this chart.*\n")
}

# ── KPIs ──────────────────────────────────────────────────────
fmt <- function(x) if (is.na(x) || length(x) == 0) "—" else comma(x)

kpi_total <- if (data_loaded) sum(instruction_attendance$attendance_count) else NA_integer_
kpi_ucla  <- if (data_loaded) {
  sum(instruction_attendance$attendance_count[instruction_attendance$institution == "UCLA"], na.rm = TRUE)
} else NA_integer_
kpi_unknown_affiliation <- if (data_loaded) {
  sum(instruction_attendance$attendance_count[is.na(instruction_attendance$institution)])
} else NA_integer_
kpi_organizational_parent <- if (data_loaded) {
  sum(instruction_attendance$attendance_count[
    instruction_attendance$institution == "UCLA" &
      !is.na(instruction_attendance$organizational_parent)
  ], na.rm = TRUE)
} else NA_integer_
kpi_years <- if (data_loaded) n_distinct(year(instruction_attendance$date), na.rm = TRUE) else NA_integer_

# ── Date range label and YoY comparison ───────────────────────
data_range_label <- if (data_loaded) {
  yr_min <- min(year(instruction_attendance$date), na.rm = TRUE)
  yr_max <- max(year(instruction_attendance$date), na.rm = TRUE)
  if (yr_min == yr_max) as.character(yr_min) else paste0(yr_min, "\u2013", yr_max)
} else "All Time"

yoy_attendance <- if (data_loaded) {
  by_year <- instruction_attendance |>
    filter(!is.na(date)) |>
    mutate(year = year(date)) |>
    summarise(n = sum(attendance_count), .by = year) |>
    arrange(desc(year)) |>
    head(2)
  if (nrow(by_year) >= 2) {
    n_recent <- by_year$n[1];  yr_recent <- by_year$year[1]
    n_prior  <- by_year$n[2];  yr_prior  <- by_year$year[2]
    pct   <- round((n_recent - n_prior) / n_prior * 100)
    arrow <- if (pct > 0) "\u2191" else if (pct < 0) "\u2193" else "\u2192"
    sprintf("%s %+d%% \u00b7 %s in %d vs %s in %d",
            arrow, pct, comma(n_recent), yr_recent, comma(n_prior), yr_prior)
  } else ""
} else ""

# ── Pre-computed alt text ─────────────────────────────────────
if (data_loaded) {
  .yr <- instruction_attendance |> filter(!is.na(date)) |>
    mutate(Year = year(date)) |>
    summarise(n = sum(attendance_count), .by = Year) |> arrange(Year)
  alt_year <- sprintf(
    "Horizontal bar chart: workshop attendance by year from %d to %d. %s.",
    min(.yr$Year), max(.yr$Year),
    paste(.yr$Year, .yr$n, sep = ": ", collapse = "; ")
  )
  .dept <- instruction_attendance |>
    filter(!is.na(standardized_department), institution == "UCLA") |>
    summarise(n = sum(attendance_count), .by = standardized_department) |>
    arrange(desc(n)) |> head(15)
  alt_dept <- if (nrow(.dept) > 0) sprintf(
      "Horizontal bar chart: top %d UCLA departments by workshop attendance. Highest: %s (%d).",
      nrow(.dept), .dept$standardized_department[1], .dept$n[1]
    ) else "Workshop attendance by UCLA department (no known departments)"
  .affil <- instruction_attendance |> filter(!is.na(status)) |>
    summarise(n = sum(attendance_count), .by = status) |>
    arrange(desc(n)) |> head(5)
  alt_affil <- if (nrow(.affil) > 0) sprintf(
      "Horizontal bar chart: workshop attendance by participant affiliation. Top group: %s (%d).",
      .affil$status[1], .affil$n[1]
    ) else "Workshop attendance by affiliation (no known affiliations)"
} else {
  alt_year  <- "Workshop attendance by year (data not loaded)"
  alt_dept  <- "Top departments by workshop attendance (data not loaded)"
  alt_affil <- "Attendance by affiliation (data not loaded)"
}
```

*For DSC staff and library administration — workshop attendance from attendee-level and validated aggregate sources.*

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*

## At a Glance (`r data_range_label`)

::: {.kpi-row .grid}

::: {.g-col-6 .g-col-md-4 .kpi-card}
[Total Attendance Records]{.kpi-label}
[`r fmt(kpi_total)`]{.kpi-value}
[All workshop participants, all years`r if (nchar(yoy_attendance) > 0) paste0(" \u00b7 ", yoy_attendance)`]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[UCLA-Affiliated]{.kpi-label}
[`r fmt(kpi_ucla)`]{.kpi-value}
[institution == "UCLA"]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[Years with Data]{.kpi-label}
[`r fmt(kpi_years)`]{.kpi-value}
[Distinct calendar years]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[UCLA Affiliation Unknown]{.kpi-label}
[`r fmt(kpi_unknown_affiliation)`]{.kpi-value}
[Included in totals; excluded from affiliation, department, and school metrics]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[UCLA Organizational Parent Known]{.kpi-label}
[`r fmt(kpi_organizational_parent)`]{.kpi-value}
[`r percent(kpi_organizational_parent / kpi_ucla, accuracy = 0.1)` of validated UCLA attendance]{.kpi-note}
:::

:::

## Workshop Attendance

::: {.panel-tabset}

### By Year

```{r chart_year, echo=FALSE, fig.alt=alt_year}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(date)) |>
    mutate(Year = year(date)) |>
    summarise(n = sum(attendance_count), .by = Year) |>
    ggplot(aes(x = factor(Year), y = n)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    labs(x = NULL, y = "Attendance") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

### UC-Wide Series

```{r uc_series_table, echo=FALSE}
if (data_loaded) {
  instruction_attendance |>
    mutate(
      Series = case_when(
        str_detect(str_to_lower(coalesce(event, "")), "love data|ldw") ~ "Love Data Week",
        str_detect(str_to_lower(coalesce(event, "")), "gis week|uc gis") ~ "GIS Week",
        TRUE ~ NA_character_
      ),
      Year = year(date)
    ) |>
    filter(!is.na(Series), !is.na(Year)) |>
    summarise(Attendance = sum(attendance_count), .by = c(Series, Year)) |>
    arrange(Series, Year) |>
    knitr::kable(
      align = c("l", "r", "r"),
      caption = "Attendance instances in the consolidated UC-wide workshop series"
    )
}
```

### By Department

```{r chart_dept, echo=FALSE, fig.alt=alt_dept}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(standardized_department), institution == "UCLA") |>
    summarise(n = sum(attendance_count), .by = standardized_department) |>
    arrange(desc(n)) |>
    head(15) |>
    ggplot(aes(x = reorder(standardized_department, n), y = n)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    labs(x = NULL, y = "Attendance", title = "Top 15 Departments (UCLA attendees)") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

```{r dept_table, echo=FALSE}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(standardized_department), institution == "UCLA") |>
    summarise(n = sum(attendance_count), .by = standardized_department) |>
    arrange(desc(n)) |>
    rename(Department = standardized_department, Attendance = n) |>
    DT::datatable(
      rownames = FALSE,
      options  = list(pageLength = 10, dom = "ftp"),
      caption  = "All UCLA departments — search or sort to find a specific unit"
    )
}
```

### By Organizational Parent

```{r organizational_parent_table, echo=FALSE}
if (data_loaded) {
  instruction_attendance |>
    filter(institution == "UCLA", !is.na(organizational_parent)) |>
    summarise(Attendance = sum(attendance_count), .by = c(organizational_parent, parent_type)) |>
    arrange(desc(Attendance)) |>
    rename(`Organizational parent` = organizational_parent, `Parent type` = parent_type) |>
    DT::datatable(
      rownames = FALSE,
      options = list(pageLength = 15, dom = "ftp"),
      caption = "Validated UCLA attendance by broader organizational parent"
    )
}
```

### By Affiliation

```{r chart_affil, echo=FALSE, fig.alt=alt_affil}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(status)) |>
    summarise(n = sum(attendance_count), .by = status) |>
    arrange(desc(n)) |>
    head(5) |>
    ggplot(aes(x = reorder(status, n), y = n)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    labs(x = NULL, y = "Attendance") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

```{r affil_table, echo=FALSE}
if (data_loaded) {
  instruction_attendance |>
    filter(!is.na(status)) |>
    summarise(n = sum(attendance_count), .by = status) |>
    arrange(desc(n)) |>
    rename(Affiliation = status, Attendance = n) |>
    DT::datatable(
      rownames = FALSE,
      options  = list(pageLength = 10, dom = "ftp"),
      caption  = "All participant affiliations"
    )
}
```

:::

**See also:** [Consulting <span aria-hidden="true">→</span>](consulting.qmd) one-on-one consultations · [Infrastructure <span aria-hidden="true">→</span>](infrastructure.qmd) Dataverse and AWS metrics · [About & methodology](about.qmd)

---

::: {.callout-note collapse="true" .data-notes title="Data notes"}
- **Sources**: Consolidated workshop attendee records (including UC-wide GIS Week and Love Data Week) and privacy-safe aggregate workshop summaries
- **Unit**: Attendance instances. LibInsight uses one attendee-event row; aggregate sources provide a validated count for each event. A person attending three workshops contributes three attendance instances
- **UCLA attendees**: filtered on `institution == "UCLA"`. Records with unknown institution remain in overall/year/topic totals but are excluded from UCLA, department, and school metrics
- **Department**: `standardized_department` field — mapped from raw free-text entries; unmapped values appear as-is
- **Organizational parent**: versioned mapping from validated UCLA departments to College divisions, professional schools, the Library, UCLA Health, institutes, or administrative groups. Missing and unresolved departments remain unknown
- **School/division**: not inferred when the source lacks a validated department. Unknown is never treated as non-UCLA
- **UC-wide series**: GIS Week and Love Data Week are included from the consolidated attendee source; their upstream workbooks are retained as provenance and are not added again
- Full rules: [Metrics Governance](metrics-governance.md) · [About & methodology](about.qmd)
:::

```{r coverage, echo=FALSE, results='asis'}
if (data_loaded && !is.null(instruction_attendance$date)) {
  instruction_attendance |>
    mutate(`Record type` = if_else(record_granularity == "aggregate", "event_summary", "attendee_event")) |>
    summarise(
      Rows = n(),
      Attendance = sum(attendance_count),
      `Date range` = paste(format(min(date, na.rm = TRUE), "%Y-%m-%d"), "\u2013",
                           format(max(date, na.rm = TRUE), "%Y-%m-%d")),
      .by = c(source_system, `Record type`)
    ) |>
    rename(Source = source_system) |>
    mutate(Rows = scales::comma(Rows), Attendance = scales::comma(Attendance)) |>
    knitr::kable(align = c("l", "l", "r", "r", "l"))
}
```

::: {.callout-note collapse="true" title="Cite this page"}
UCLA Library Data Science Center. (`r year(Sys.Date())`). *Instruction Statistics*. Retrieved from <https://ucla-data-science-center.github.io/dsc-stats-reports/instruction.html>
:::

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
