---
title: "Instruction Statistics"
format:
  html:
    toc: true
    code-fold: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)

# ── Workshop data (graceful fallback) ─────────────────────────
data_loaded    <- FALSE
ucla_workshops <- tibble()

for (p in c(
  "data/processed/instruction/ucla_workshops.rda",
  "data/instruction/ucla_workshops.rda",
  "data/consultations/Data/ucla_workshops.rda"
)) {
  if (file.exists(p)) {
    load(p)
    data_loaded <- TRUE
    break
  }
}

if (!data_loaded) {
  message("ucla_workshops.rda not found — charts will show placeholder text.")
}

no_data_msg <- function() {
  cat("*Data file not loaded. Run the ETL scripts to populate this chart.*\n")
}

# ── KPIs ──────────────────────────────────────────────────────
fmt <- function(x) if (is.na(x) || length(x) == 0) "—" else comma(x)

kpi_total <- if (data_loaded) nrow(ucla_workshops)                                       else NA_integer_
kpi_ucla  <- if (data_loaded) sum(ucla_workshops$institution == "UCLA", na.rm = TRUE)    else NA_integer_
kpi_years <- if (data_loaded) n_distinct(year(ucla_workshops$date), na.rm = TRUE)        else NA_integer_

# ── Pre-computed alt text ─────────────────────────────────────
if (data_loaded) {
  .yr <- ucla_workshops |> filter(!is.na(date)) |>
    mutate(Year = year(date)) |> count(Year) |> arrange(Year)
  alt_year <- sprintf(
    "Horizontal bar chart: workshop attendance by year from %d to %d. %s.",
    min(.yr$Year), max(.yr$Year),
    paste(.yr$Year, .yr$n, sep = ": ", collapse = "; ")
  )
  .dept <- ucla_workshops |>
    filter(!is.na(standardized_department), institution == "UCLA") |>
    count(standardized_department, sort = TRUE) |> head(15)
  alt_dept <- sprintf(
    "Horizontal bar chart: top %d UCLA departments by workshop attendance. Highest: %s (%d).",
    nrow(.dept), .dept$standardized_department[1], .dept$n[1]
  )
  .affil <- ucla_workshops |> filter(!is.na(status)) |>
    count(status, sort = TRUE) |> head(5)
  alt_affil <- sprintf(
    "Horizontal bar chart: workshop attendance by participant affiliation. Top group: %s (%d).",
    .affil$status[1], .affil$n[1]
  )
} else {
  alt_year  <- "Workshop attendance by year (data not loaded)"
  alt_dept  <- "Top departments by workshop attendance (data not loaded)"
  alt_affil <- "Attendance by affiliation (data not loaded)"
}
```

## At a Glance

::: {.kpi-row .grid}

::: {.g-col-6 .g-col-md-4 .kpi-card}
[Total Attendance Records]{.kpi-label}
[`r fmt(kpi_total)`]{.kpi-value}
[All workshop participants, all years]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[UCLA-Affiliated]{.kpi-label}
[`r fmt(kpi_ucla)`]{.kpi-value}
[institution == "UCLA"]{.kpi-note}
:::

::: {.g-col-6 .g-col-md-4 .kpi-card}
[Years with Data]{.kpi-label}
[`r fmt(kpi_years)`]{.kpi-value}
[Distinct calendar years]{.kpi-note}
:::

:::

## Workshop Attendance

::: {.panel-tabset}

### By Year

```{r chart_year, echo=FALSE, fig.alt=alt_year}
if (data_loaded) {
  ucla_workshops |>
    filter(!is.na(date)) |>
    mutate(Year = year(date)) |>
    count(Year) |>
    ggplot(aes(x = factor(Year), y = n)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    labs(x = NULL, y = "Attendance") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

### By Department

```{r chart_dept, echo=FALSE, fig.alt=alt_dept}
if (data_loaded) {
  ucla_workshops |>
    filter(!is.na(standardized_department), institution == "UCLA") |>
    count(standardized_department, sort = TRUE) |>
    head(15) |>
    ggplot(aes(x = reorder(standardized_department, n), y = n)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    labs(x = NULL, y = "Attendance", title = "Top 15 Departments (UCLA attendees)") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

### By Affiliation

```{r chart_affil, echo=FALSE, fig.alt=alt_affil}
if (data_loaded) {
  ucla_workshops |>
    filter(!is.na(status)) |>
    count(status, sort = TRUE) |>
    head(5) |>
    ggplot(aes(x = reorder(status, n), y = n)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    labs(x = NULL, y = "Attendance") +
    theme_minimal() +
    geom_text(aes(label = n), color = "white", hjust = 1.2)
} else { no_data_msg() }
```

:::

---

::: {.callout-note collapse="true" .data-notes title="Data notes"}
- **Source**: LibInsight workshop/event attendance exports
- **Unit**: One row = one participant registration. A person attending three workshops is counted three times — totals reflect attendee-events, not unique individuals
- **UCLA attendees**: filtered on `institution == "UCLA"`; external attendees included in totals but excluded from department breakdowns
- **Department**: `standardized_department` field — mapped from raw free-text entries; unmapped values appear as-is
- Full rules: [Metrics Governance](metrics-governance.md) · [About](about.qmd)
:::

```{r coverage, echo=FALSE, results='asis'}
if (data_loaded && !is.null(ucla_workshops$date)) {
  tibble(
    Source        = "LibInsight",
    `Record type` = "workshop_attendance",
    Rows          = scales::comma(nrow(ucla_workshops)),
    `Date range`  = paste(
      format(min(ucla_workshops$date, na.rm = TRUE), "%Y-%m-%d"), "\u2013",
      format(max(ucla_workshops$date, na.rm = TRUE), "%Y-%m-%d")
    )
  ) |>
    knitr::kable(align = c("l", "l", "r", "l"))
}
```

::: {.callout-note collapse="true" title="Cite this page"}
UCLA Library Data Science Center. (`r year(Sys.Date())`). *Instruction Statistics*. Retrieved from <https://ucla-data-science-center.github.io/dsc-stats-reports/instruction.html>
:::

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
