---
title: "Research Enabled"
subtitle: "Pathway: one-on-one research data support at UCLA, and the capacity behind it"
toc: true
toc-depth: 2
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)

read_if <- function(path) if (file.exists(path)) read_csv(path, show_col_types = FALSE) else tibble()
headline <- read_if("data/processed/canonical/headline_aggregates.csv")
hl <- function(key) {
  x <- headline$display[headline$metric_id == key]
  if (length(x) == 0) "—" else x[[1]]
}
series   <- read_if("data/processed/canonical/consultations_by_year.csv")
capacity <- read_if("data/processed/canonical/service_capacity_by_year.csv")

alt_combined <- if (nrow(series) > 0 && nrow(capacity) > 0) sprintf(
  "Three horizontal bar charts on the same year rows, 2017 to 2026. Recorded consultations: %s (2024 onward are floors). Staff time in FTE: %s. Average students on payroll per month: %s.",
  paste(series$year, series$display, sep = ": ", collapse = "; "),
  paste(capacity$year, coalesce(as.character(capacity$dsc_staff_fte + capacity$contributed_staff_fte), "no record"), sep = ": ", collapse = "; "),
  paste(capacity$year, coalesce(as.character(capacity$student_avg_monthly), "no record"), sep = ": ", collapse = "; ")) else "Recorded consultations and service capacity by year (data not loaded)"
```

[]{.brand-el .brand-node1 .page-mark aria-hidden="true"}

## Finding and scope

DSC's staff time fell from 5.7 FTE in 2022-23 to 4.0 in 2025-26, and student help fell by half in 2025, while consultation records became incomplete. Researchers across UCLA bring data problems to DSC, and some of those consultations become multi-year collaborations. DSC recorded `r hl("consultations_recorded_series")` scheduled consultations from 2017 to September 2026. Recorded counts rise and fall with the staff and students available to do the work and with the system used to record it, so they show the work delivered, not the demand for it.

## Evidence in brief

The figure shows capacity and recorded delivery side by side, not an estimate of demand.

```{r evidence_chart, echo=FALSE, fig.alt=alt_combined, fig.width=10, fig.height=5, fig.cap="Left: recorded consultations (scheduled appointments with DSC staff and DataSquad consultants); from mid-2024 recording lapsed (Calendly ended; LibCal from May 2025), so 2024 onward are floors, and 2026 runs to September. Middle: staff time in full-time equivalents, DSC staff plus time contributed by other Library units (one share shown at its upper bound); staff time covers all DSC services, not consultations only. Right: average student employees on payroll per month (UC timesheet records; 2026 through July); from July to October 2025, while student hiring was under campus review, one student remained on payroll."}
if (nrow(series) > 0 && nrow(capacity) > 0) {
  yrs <- as.character(rev(sort(union(series$year, capacity$year))))
  panels <- c("Recorded consultations", "Staff (FTE)", "Students on payroll")
  cons <- series |> transmute(year = as.character(year), panel = panels[1], value,
                              series = if_else(is_floor, "Floor (recording incomplete)", "Full recording"), label = display)
  staff <- capacity |>
    transmute(year = as.character(year), `DSC staff` = dsc_staff_fte, `Other units` = contributed_staff_fte) |>
    pivot_longer(-year, names_to = "series", values_to = "value") |> mutate(panel = panels[2], label = NA_character_)
  stud <- capacity |> transmute(year = as.character(year), panel = panels[3], value = student_avg_monthly,
                                series = "Students", label = NA_character_)
  bars <- bind_rows(cons, staff, stud)
  totals <- bind_rows(
    cons |> select(year, panel, value, label),
    staff |> group_by(year, panel) |>
      summarise(value = if (all(is.na(value))) NA_real_ else sum(value, na.rm = TRUE), .groups = "drop") |>
      mutate(label = if_else(is.na(value), "no record", number(value, accuracy = 0.1))),
    stud |> mutate(label = if_else(is.na(value), "no record", number(value, accuracy = 0.1))) |> select(year, panel, value, label))
  fills <- c("Full recording" = "#2774AE", "Floor (recording incomplete)" = "#A9C7E3",
             "DSC staff" = "#003B5C", "Other units" = "#8BB8E8", "Students" = "#FFB81C")
  bars |>
    mutate(year = factor(year, levels = yrs), panel = factor(panel, levels = panels),
           series = factor(series, levels = c("Other units", "DSC staff", names(fills)[c(1, 2, 5)]))) |>
    ggplot(aes(y = year, x = coalesce(value, 0), fill = series)) +
    geom_col(width = 0.7) +
    geom_text(data = totals |> mutate(year = factor(year, levels = yrs), panel = factor(panel, levels = panels)),
              aes(y = year, x = coalesce(value, 0), label = label), inherit.aes = FALSE, hjust = -0.15, size = 4) +
    facet_wrap(~panel, scales = "free_x", nrow = 1) +
    scale_y_discrete(drop = FALSE) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.28)), n.breaks = 4) +
    scale_fill_manual(values = fills, breaks = names(fills), name = NULL) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 14) +
    theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
          legend.position = "bottom", strip.text = element_text(face = "bold", size = 14),
          plot.margin = margin(5, 18, 5, 5), panel.spacing.x = unit(2.2, "lines")) +
    guides(fill = guide_legend(nrow = 1))
}
```

Recorded counts from mid-2024 on are floors: the scheduling system lapsed while staff and student capacity fell, so these years cannot show whether demand changed. Consultations are interactions, not people.

::: {.callout-note collapse="true" title="How the consultation series is counted"}
- About 340 to 395 consultations a year were recorded at full recording (2021-2023), some growing into multi-year collaborations.
- From mid-2024 the scheduling system lapsed (Calendly ended; LibCal from May 2025), so later counts are floors and cannot separate demand from the capacity to meet it or the ability to record it.
- Consultations are interactions, not people. One researcher can book many appointments; the separate 1,908 figure counts people.
- The series counts scheduled appointments, with DSC staff and DataSquad consultants alike. DataSquad's in-person walk-in sign-ins are shown separately on the [consultation evidence page](consulting.qmd).
:::

Two supporting measures:

- **At least `r hl("people_reached_confirmed_ucla")` confirmed UCLA researchers** used DSC instruction or consultation, 2017 to September 2026 (a floor, counted only where identity could be confirmed).
- **Scheduling systems changed three times:** Calendly to mid-2024, no scheduling system until May 2025, then LibCal. Consultations continued in the gap, but most went unrecorded.

## A research collaboration: making helicopter-surveillance data usable

The Carceral Ecologies lab in UCLA's Institute for Society and Genetics needed to analyze a large dataset of law-enforcement helicopter flights over Los Angeles, again and again. Over several years, DSC and DataSquad provided an R server, helped build the processing workflow, and trained a student researcher to keep using it.

**One processing task went from about 5 hours 18 minutes to about 2 minutes.** That made repeated analysis of the dataset practical for the lab. It is a result from this task, not a benchmark for DSC's work in general.

> "This work wouldn't have happened without your team!"
>
> Nicholas Shapiro, director of Carceral Ecologies, November 2024 (quoted with permission)

The research and its findings belong to the lab. DSC's documented contribution was the computing, the workflow, and the training that supported it (claim #1). The lab's work was later recognized with a 2024 UCLA Public Impact Research Award.

::: {.callout-note .long-view}
## The long view (history, as of 2026)

The archive answered research questions from its early years: data discovery, codebooks, and help with statistical software. The 2018 redesign widened consulting to the whole research life cycle, from data management plans to publication and preservation, and added GIS, 3D capture, machine learning, research software, and reproducible workflows. From about 2023 the work shifted from single appointments toward project support that ran for months. [Read the consulting section of the history](https://ucla-data-science-center.github.io/ssda-dsc-history/history.html#consulting-from-reference-desk-to-research-partner)
:::

## Detailed evidence

The consultation evidence page holds the full tables. Jump to:

- [Recorded consultations by year](consulting.qmd#recorded-consultations-since-2017)
- [Consultation records by record type, department, and affiliation](consulting.qmd#direct-consultations) (records file, January 2021 to June 2024; breakdowns suppress any unit under 5)
- [Research impact cases](consulting.qmd#research-impact)
- [Methodology](about.qmd)

## Sources and claim record

| Figure | Window | Unit | Extracted | Claim |
|---|---|---|---|---|
| `r hl("consultations_recorded_series")` recorded consultations | 2017 to Sept 2026 | scheduled appointment | 2026-09-27 | #10 |
| At least `r hl("people_reached_confirmed_ucla")` UCLA researchers | 2017 to Sept 2026 | person, confirmed | 2026-09-26 | #13 |
| 318 to 2 minute processing task | one engagement | one task | 2026-09-12 | #1 |
| Student employees on payroll | 2018 to Jul 2026 | monthly average headcount | 2026-09-27 | #17 |
| Consulting history | 1961 to 2026 | historical account | 2026-09-27 | #18 |

Suppression rule: department, organizational-parent, and affiliation breakdowns fold any unit under 5 into one combined row, which is omitted if it is also under 5. Approved wording and sources: [claim-register.md](https://github.com/ucla-data-science-center/dsc-stats-reports/blob/main/claim-register.md).

::: {.callout-note collapse="true" title="Cite this page"}
UCLA Library Data Science Center. (`r format(Sys.Date(), "%Y")`). *Research Enabled*. Retrieved from <https://ucla-data-science-center.github.io/dsc-stats-reports/research.html>
:::

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
