---
title: "Research Enabled"
subtitle: "Pathway: one-on-one research data support at UCLA, and the capacity behind it"
toc: true
toc-depth: 2
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)

read_if <- function(path) if (file.exists(path)) read_csv(path, show_col_types = FALSE) else tibble()
headline <- read_if("data/processed/canonical/headline_aggregates.csv")
hl <- function(key) {
  x <- headline$display[headline$metric_id == key]
  if (length(x) == 0) "—" else x[[1]]
}
series   <- read_if("data/processed/canonical/consultations_by_year.csv")
capacity <- read_if("data/processed/canonical/service_capacity_by_year.csv")

alt_combined <- if (nrow(series) > 0 && nrow(capacity) > 0) sprintf(
  "Three horizontal bar charts on the same year rows, 2017 to 2026. Recorded consultations: %s (2024 onward are floors). Staff time in FTE: %s. Average students on payroll per month: %s.",
  paste(series$year, series$display, sep = ": ", collapse = "; "),
  paste(capacity$year, coalesce(as.character(capacity$dsc_staff_fte + capacity$contributed_staff_fte), "no record"), sep = ": ", collapse = "; "),
  paste(capacity$year, coalesce(as.character(capacity$student_avg_monthly), "no record"), sep = ": ", collapse = "; ")) else "Recorded consultations and service capacity by year (data not loaded)"
```

[]{.brand-el .brand-node1 .page-mark aria-hidden="true"}

## Finding and scope

DSC's staff time fell from 5.7 FTE in 2022-23 to 4.0 in 2025-26, and student help fell by half in 2025, while consultation records became incomplete. Researchers across UCLA bring data problems to DSC, and some of those consultations become multi-year collaborations. DSC recorded `r hl("consultations_recorded_series")` scheduled consultations from 2017 to September 2026. Recorded counts rise and fall with the staff and students available to do the work and with the system used to record it, so they show the work delivered, not the demand for it.

## Evidence in brief

The figure shows capacity and recorded delivery side by side, not an estimate of demand.

```{r evidence_chart, echo=FALSE, fig.alt=alt_combined, fig.width=10, fig.height=5, fig.cap="Left: recorded consultations (scheduled appointments with DSC staff and DataSquad consultants); from mid-2024 recording lapsed (Calendly ended; LibCal from May 2025), so 2024 onward are floors, and 2026 runs to September. Middle: staff time in full-time equivalents, DSC staff plus time contributed by other Library units (one share shown at its upper bound); staff time covers all DSC services, not consultations only. Right: average student employees on payroll per month (UC timesheet records; 2026 through July); from July to October 2025, while student hiring was under campus review, one student remained on payroll."}
if (nrow(series) > 0 && nrow(capacity) > 0) {
  yrs <- as.character(rev(sort(union(series$year, capacity$year))))
  panels <- c("Recorded consultations", "Staff (FTE)", "Students on payroll")
  cons <- series |> transmute(year = as.character(year), panel = panels[1], value,
                              series = if_else(is_floor, "Floor (recording incomplete)", "Full recording"), label = display)
  staff <- capacity |>
    transmute(year = as.character(year), `DSC staff` = dsc_staff_fte, `Other units` = contributed_staff_fte) |>
    pivot_longer(-year, names_to = "series", values_to = "value") |> mutate(panel = panels[2], label = NA_character_)
  stud <- capacity |> transmute(year = as.character(year), panel = panels[3], value = student_avg_monthly,
                                series = "Students", label = NA_character_)
  bars <- bind_rows(cons, staff, stud)
  totals <- bind_rows(
    cons |> select(year, panel, value, label),
    staff |> group_by(year, panel) |>
      summarise(value = if (all(is.na(value))) NA_real_ else sum(value, na.rm = TRUE), .groups = "drop") |>
      mutate(label = if_else(is.na(value), "no record", number(value, accuracy = 0.1))),
    stud |> mutate(label = if_else(is.na(value), "no record", number(value, accuracy = 0.1))) |> select(year, panel, value, label))
  fills <- c("Full recording" = "#2774AE", "Floor (recording incomplete)" = "#A9C7E3",
             "DSC staff" = "#003B5C", "Other units" = "#8BB8E8", "Students" = "#FFB81C")
  bars |>
    mutate(year = factor(year, levels = yrs), panel = factor(panel, levels = panels),
           series = factor(series, levels = c("Other units", "DSC staff", names(fills)[c(1, 2, 5)]))) |>
    ggplot(aes(y = year, x = coalesce(value, 0), fill = series)) +
    geom_col(width = 0.7) +
    geom_text(data = totals |> mutate(year = factor(year, levels = yrs), panel = factor(panel, levels = panels)),
              aes(y = year, x = coalesce(value, 0), label = label), inherit.aes = FALSE, hjust = -0.15, size = 4) +
    facet_wrap(~panel, scales = "free_x", nrow = 1) +
    scale_y_discrete(drop = FALSE) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.28)), n.breaks = 4) +
    scale_fill_manual(values = fills, breaks = names(fills), name = NULL) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 14) +
    theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
          legend.position = "bottom", strip.text = element_text(face = "bold", size = 14),
          plot.margin = margin(5, 18, 5, 5), panel.spacing.x = unit(2.2, "lines")) +
    guides(fill = guide_legend(nrow = 1))
}
```

Recorded counts from mid-2024 on are floors: the scheduling system lapsed while staff and student capacity fell, so these years cannot show whether demand changed. Consultations are interactions, not people.

::: {.callout-note collapse="true" title="How the consultation series is counted"}
- About 340 to 395 consultations a year were recorded at full recording (2021-2023), some growing into multi-year collaborations.
- From mid-2024 the scheduling system lapsed (Calendly ended; LibCal from May 2025), so later counts are floors and cannot separate demand from the capacity to meet it or the ability to record it.
- Consultations are interactions, not people. One researcher can book many appointments; the separate 1,908 figure counts people.
- The series counts scheduled appointments, with DSC staff and DataSquad consultants alike. DataSquad's in-person walk-in sign-ins are shown separately on the [consultation evidence page](consulting.qmd).
:::

Two supporting measures:

- **At least `r hl("people_reached_confirmed_ucla")` confirmed UCLA researchers** used DSC instruction or consultation, 2017 to September 2026 (a floor, counted only where identity could be confirmed).
- **Scheduling systems changed three times:** Calendly to mid-2024, no scheduling system until May 2025, then LibCal. Consultations continued in the gap, but most went unrecorded.

## A research collaboration: LAPD helicopter surveillance

The Carceral Ecologies lab, directed by Nicholas Shapiro in UCLA's Institute for Society and Genetics, studies the health and mental-health effects of law-enforcement helicopter surveillance across Los Angeles County from 2017 through 2022 ([lab project page](https://carceralecologies.org/projects/)). Its public findings include that LAPD helicopters fly lowest over largely Black and Hispanic neighborhoods, with a median of about 550 feet over Watts at night ([CalMatters, July 2024](https://calmatters.org/commentary/2024/07/california-longstanding-bias-police-helicopters/)).

A student researcher on the team first learned R and GIS from DSC staff in a 2019 African American Studies course taught with DSC. In 2022 she came back with the team's flight-tracking data, twelve monthly files of about a million rows each, and met DSC staff and DataSquad consultants more than 30 times that year ([her 2023 interview](https://ucla-datasquad.github.io/2023/07/27/kate-m-helicopter-data.html)). Together they converted the raw flight records into analyzable tables, identified individual flights, filtered out impossible altitude readings, and joined millions of positions to Los Angeles neighborhoods; DSC's GIS librarian advised on the spatial and statistical analysis ([her analysis code](https://github.com/kmcinerny/LA_Law_Enforcement_Heli)).

As the lab acquired more data, DSC helped it scale. In 2023 DSC staff and DataSquad students worked with the lab's statistician to process the dataset in batches, speed up elevation lookups about tenfold by switching data sources, and run long jobs on DSC's deep learning machine and hosted R server ([shared code](https://github.com/ucla-data-science-center/helicopter-data)). One measured piece of that work cut the time to load the data from about 318 minutes to about 2, about 159 times as fast ([DataSquad project page](https://ucla-datasquad.github.io/)). DSC staff on the collaboration included Tim Dennis, Leigh Phan, Zhiyuan Yao, and Kristian Allen.

> "This work wouldn't have happened without your team!"
>
> Nicholas Shapiro, director of Carceral Ecologies, November 2024 (quoted with permission)

The research, its findings, and the public attention it has drawn belong to the lab. DSC's part was the data work, computing, and training behind it. The lab's work was recognized with a 2024 UCLA Public Impact Research Award.

::: {.callout-note .long-view}
## The long view (history, as of 2026)

The archive answered research questions from its early years: data discovery, codebooks, and help with statistical software. The 2018 redesign widened consulting to the whole research life cycle, from data management plans to publication and preservation, and added GIS, 3D capture, machine learning, research software, and reproducible workflows. From about 2023 the work shifted from single appointments toward project support that ran for months. [Read the consulting section of the history](https://ucla-data-science-center.github.io/ssda-dsc-history/history.html#consulting-from-reference-desk-to-research-partner)
:::

## Detailed evidence

The consultation evidence page holds the full tables. Jump to:

- [Recorded consultations by year](consulting.qmd#recorded-consultations-since-2017)
- [Consultation records by record type, department, and affiliation](consulting.qmd#direct-consultations) (records file, January 2021 to June 2024; breakdowns suppress any unit under 5)
- [Research impact cases](consulting.qmd#research-impact)
- [Methodology](about.qmd)

## Sources and claim record

| Figure | Window | Unit | Extracted | Claim |
|---|---|---|---|---|
| `r hl("consultations_recorded_series")` recorded consultations | 2017 to Sept 2026 | scheduled appointment | 2026-09-27 | #10 |
| At least `r hl("people_reached_confirmed_ucla")` UCLA researchers | 2017 to Sept 2026 | person, confirmed | 2026-09-26 | #13 |
| Carceral Ecologies collaboration; 318 to 2 minute data-loading step | 2019 to 2025 | one engagement | 2026-09-27 | #1 |
| Student employees on payroll | 2018 to Jul 2026 | monthly average headcount | 2026-09-27 | #17 |
| Consulting history | 1961 to 2026 | historical account | 2026-09-27 | #18 |

Suppression rule: department, organizational-parent, and affiliation breakdowns fold any unit under 5 into one combined row, which is omitted if it is also under 5. Approved wording and sources: [claim-register.md](https://github.com/ucla-data-science-center/dsc-stats-reports/blob/main/claim-register.md).

::: {.callout-note collapse="true" title="Cite this page"}
UCLA Library Data Science Center. (`r format(Sys.Date(), "%Y")`). *Research Enabled*. Retrieved from <https://ucla-data-science-center.github.io/dsc-stats-reports/research.html>
:::

*Last rendered: `r format(Sys.Date(), "%B %d, %Y")`*
