---
title: "Data Science Center - FY 2018/2019 Statistics"
output: 
  beamer_presentation:
    theme: ucla
    latex_engine: xelatex
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=F, message=FALSE}
#install.packages("googledrive")
library(googledrive)
library(tidyverse)
library(scales)
options(stringsAsFactors = F)
library(readr)
library(RColorBrewer)
library(viridis)
library(kableExtra)
library(tidyr)
library(htmlTable)
library(stringr)
library(janitor)
library(lubridate)
library(calecopal)
```

```{r read-data, include=FALSE, message=FALSE}
#dscstats <- read_csv('data/dsc_stats_clean.csv') # cleaned via openrefine
dscstats <- read_csv('data/workshops_events.csv') #incl. external workshops, inst training
dscstats <- clean_names(dscstats)

load('data/workshop_obs.RData')
load('data/consulting_obs.RData')
```

```{r filter_fy18-19, echo=FALSE}
wkshop18_19 <- workshop_obs %>% 
  filter(date >= "2018-07-01" & date <= "2019-06-30")
```


```{r color_palette, echo=FALSE}
# The palette with grey:
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
uclaPalette <- c("#2774AE", "#FFD100", "#003B5C", "#00FF87", "#FFB81C", "#8237FF", "#D55E00", "#CC79A7","#2774AE")
```

## Instruction Fiscal Year 2018/19

* Addresses researcher, staff, and librarian skills gaps in data science and foundation coding skills. 
* Member of the Carpentries, a global community of volunteer instructors
* Contribute to curricula and train-the-trainer best practices nationally and international. 
* From July 2018 to June 2019, this program has offered **`r length(unique(wkshop18_19$date))`** workshops or educational events to the UCLA community.


## Workshop attendance by campus role 

```{r by-audience, echo=FALSE}
wkshop18_19 %>% 
  filter(institution == "UCLA") %>% 
  count(role) %>% 
  drop_na() %>% 
  ggplot(aes(x=reorder(str_to_title(role), -n), y=n)) + 
    geom_col(fill="#2774AE") +
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1)) + 
    labs(x = "Status", y="Attendance") + 
    geom_text (aes(label=n), hjust= -0.12) +
    theme_minimal()
```

## Attendance by data science topic by role 

```{r attendance-by-topic, echo=FALSE}
workshop_obs %>% 
  group_by(practice_topic, role) %>% 
  summarize(number_people = n()) %>% 
  drop_na(role, practice_topic) %>% 
  ggplot(aes(x=practice_topic, y=number_people, fill = str_to_title(role))) + 
    geom_col(position = "dodge") +
    coord_flip() +
    scale_fill_manual(values=cal_palette("kelp1")) +
    #scale_fill_brewer() +
    labs(title = "Attendance by Practice and Topic Taught by Audience", x= "Practice or topic", y="Attendance", fill = "Role") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    theme_minimal()
```

## Attendance by tool taught by audience


```{r Event attendance-by-tool, echo=FALSE}
workshop_obs  %>% 
  group_by(tool, role) %>% 
  summarize(number_people = n()) %>% 
  drop_na(role, tool) %>% 
  ggplot(aes(x=str_to_title(tool), y=number_people, fill = str_to_title(role))) + 
    geom_col(position = "dodge") +
    coord_flip() +
    scale_fill_manual(values=cal_palette("kelp1")) +
    #scale_fill_brewer() +
    labs(title = "Attendance by Tool Taught by Audience", x= "Data Science Tool", y="Attendance", fill = "Role") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    theme_minimal()
```

## Attendance by tool: 

```{r attendance_role_status, echo=FALSE}
#reshape to include tool2
wkshop18_19  %>% 
  #gather()
  group_by(tool) %>% 
  summarize(number_status = n()) %>% 
  drop_na(tool) %>% 
  #ggplot(aes(x=reorder(str_to_title(role), -number_status), y=number_status)) + 
  ggplot(aes(x=reorder(str_to_title(tool), -number_status), y=number_status)) + 
    geom_col(fill="#2774AE") +
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1)) + 
    labs(x = "Status", y="Attendance") + 
    geom_text (aes(label=number_status), hjust= -0.12) +
    theme_minimal()
```

## Campus Reach

```{r number-depts, echo=FALSE}
num_depts <- wkshop18_19 %>% filter(dept != "Library") %>% drop_na(dept) %>% distinct(dept) %>% nrow()
```

For our workshops in 2018/19, learners came from **`r num_depts`** different departments, schools or centers. 

## Top departments by attendance

```{r top_ten_depts, echo=FALSE}
wkshop18_19 %>% filter(dept != "Library") %>% 
  select(dept) %>% 
  drop_na() %>% 
  rename(Departments = dept) %>% 
  count(Departments) %>% 
  rename(Number = n) %>% 
  arrange(desc(Number)) %>% 
  head(5) %>% 
  ggplot(aes(x=reorder(Departments, -Number), y=Number)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    labs(x= "Depatment", y="Attendance", fill = "Role") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    theme_minimal()
```

## Consultations 

We work with researchers to help them accomplish their research goals. 


```{r consulting, eval=FALSE}
consulting_obs %>% 
  group_by(year = year(start_date_time)) %>% 
  summarize(consults = sum(n())) %>% 
  ggplot(aes(reorder(year, -consults), consults)) + geom_col(fill = "#2774AE") + 
    coord_flip() + 
    scale_fill_manual(values=cal_palette("kelp1")) +
    labs(x= "Year", y="Consults") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    theme_minimal()
```

```{r consulting_topic, eval=FALSE}

consulting_obs$response_1

```

