---
title: "DSC Report: 2024"
author: "Tim Dennis"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(googledrive)
library(tidyverse)
library(scales)
options(stringsAsFactors = F)
library(readr)
library(RColorBrewer)
library(viridis)
library(kableExtra)
library(tidyr)
library(htmlTable)
library(stringr)
library(janitor)
library(lubridate)
library(calecopal)
library(tidytext) # for NLP
library(stringr) # to deal with strings
library(wordcloud) # to render wordclouds
library(DT) # for dynamic tables
library(tidytext)
library(tm)
library(topicmodels)
```

```{r read-data, include=FALSE, message=FALSE}
# Source the R script
source("data_cleaning.R")

# Load the cleaned and merged data
dsc_consult <- readRDS("data/dsc_consult_merged.rds")

# View the first few rows of the merged data
#head(dsc_consult)
```

## Events & Workshops

The DSC puts on events for the UCLA community and for the larger UC system in collaboration with other campuses. We manage the Carpentries membership and provide community for over 10 instructors on the UCLA campus. Our workshops address researchers, staff, and librarians skills gaps in data science and foundation coding skills. We contribute to curricula and train-the-trainer best practices through a regional and global network.

To get a sense of the growth of events organized, taught & designed by the DSC community, let us look at attendance over time:

## Attendance over time

```{r workshops_year, echo=FALSE}
dsc_workshops %>% 
  count(Year = year(date), name = "Number") %>%
  #need to add aggregate data from libcal for love data week
  #should find a way to fix in data 
  mutate(Number = replace(Number, Number == 1096, 2402)) %>% 
  ggplot(aes(reorder(Year, -Year), Number)) + geom_col(fill = "#2774AE") +
    coord_flip() + 
    scale_fill_manual(values=cal_palette("kelp1")) +
    labs(x= "Year", y="Attendance", title="DSC Event Attendance by Year") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    #geom_text (aes(label=Number), hjust= -0.12) +
    geom_text(aes(label=Number), color = "white", hjust = 1.2) +
    theme_minimal()
```

## Departments, Schools & Units

```{r number-depts, echo=FALSE}
num_depts_wkshp <- dsc_workshops %>% drop_na(department) %>% filter(institution == "UC Los Angeles") %>% distinct(department)  %>% nrow()
num_depts_wkshp_20_21 <- dsc_workshops %>%  filter(date >= "2020-07-01" & date <= "2021-06-30") %>% drop_na(department) %>% filter(institution == "UC Los Angeles") %>% distinct(department) %>% nrow()
num_depts_wkshp_17_20 <- dsc_workshops %>%  filter(date >= "2017-07-01" & date <= "2020-06-30") %>% drop_na(department) %>% distinct(department) %>% nrow()
```

Since 2017, our workshops have been attended by **`r num_depts_wkshp`** different departments, schools, or centers from UCLA or other universities.

## We can look at the top departments who attend our workshops: 

```{r departmens_attendance, echo=FALSE}
dsc_workshops %>% count(department, sort = TRUE, name = "Attendance") %>% drop_na(department) %>%  head(10) %>% 
  ggplot(aes(reorder(department, -Attendance), Attendance)) +
    geom_col(fill = "#2774AE") +
    coord_flip() + 
    scale_fill_manual(values=cal_palette("kelp1")) +
    labs(x= "Department", y="Attendance", title="Top Departments by Attendance", subtitle = "2017-2024") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    geom_text(aes(label=Attendance), color = "white", hjust = 1.2) +
#    geom_text (aes(label=Attendance), hjust= -0.12) +
    theme_minimal()
```

## Affiliation of Attendee: 2017-24

A look at the status of
```{r attendee_status, echo=FALSE}
dsc_workshops %>%
  # filter(date >= "2020-07-01" & date <= "2021-06-30") %>%
  count(status, sort = TRUE, name = "Attendance") %>%
  drop_na(status) %>%
  head(5) %>%
  mutate(Percentage = Attendance / sum(Attendance) * 100) %>%
  ggplot(aes(reorder(status, -Attendance), Attendance)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    scale_fill_manual(values = cal_palette("kelp1")) +
    labs(x = "Status", y = "Attendance", title = "Attendance by Status", subtitle = "2017-2022") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1)) +
    geom_text(aes(label = paste0(round(Percentage, 1), "%")), color = "white", hjust = 1.2) +
    theme_minimal()
```

## Consultations

We work with researchers one-on-one to help them accomplish their research goals. Prior to the pandemic, we provided consultations online and in-person depending on the researchers preferences. We estimate that prior to March 2020, we provided consultation online approximately 10-15% of the time. During the pandemic, we moved our service to online only and restarted in-person consulting in 2023. Regardless of how users access our service, it has shown growth over time.

```{r consulting, echo=FALSE}
dsc_consult %>% 
  group_by(year = year(start_date_time)) %>% 
  summarize(consults = sum(n())) %>% 
  arrange(desc(year)) %>%
  ggplot(aes(reorder(year, consults), consults)) + geom_col(fill = "#2774AE") + 
  #ggplot(aes(year, -consults, consults)) + geom_col(fill = "#2774AE") + 
    coord_flip() + 
    scale_fill_manual(values=cal_palette("kelp1")) +
    labs(x= "Year", y="Consults") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    geom_text(aes(label=consults), color = "white", hjust = 1.2) +
    theme_minimal() +
    theme(
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank()   # Remove minor grid lines
  ) 
```


```{r}
# Group by year and summarize the data
summary_data <- dsc_consult %>%
  group_by(year = year(start_date_time)) %>%
  summarize(consults = n()) %>%
  arrange(desc(year))  # Sort by year in descending order

# Create the bar graph with correctly ordered years
ggplot(summary_data, aes(x = factor(year, levels = rev(unique(year))), y = consults)) +
  geom_col(fill = "#2774AE") +
  coord_flip() +  # Flip the coordinates
  labs(x = "Year", y = "Consults", title = "Consults by Year") +
  theme(axis.text.y = element_text(angle = 0, size = 10, hjust = 1)) +  # Adjust text on the y-axis since we flipped
  geom_text(aes(label = consults), hjust = -0.1, color = "white") +  # Adjust text position and color
  theme_minimal()
```

## Campus Reach

```{r number-depts-consult, echo=FALSE}
num_depts <- dsc_consult %>% drop_na(department) %>% distinct(department) %>% nrow()
num_depts_20_21 <- dsc_consult %>% filter(year(start_date_time) == 2020) %>% drop_na(department) %>% distinct(department) %>% nrow()
```

From 2019, when we started to capture more user information, our consults have come from **`r num_depts`** different departments, schools or centers. We include **The Library** in this number because, as a secondary and sometimes tertiary referral point, we often work with liaising Librarians and internal units working with researchers or needing data support. For FY 2020-21, we provided data services to **`r num_depts_20_21`** campus departments.

## Consulting by Department

It is important to note that historically, we did not ask patrons to provide their departmental information in our appointment scheduler, so the data may not be fully complete. However, the data we do have indicates that the DSC service is interdisciplinary in scope and practice. This aligns with our service vision and origin, which is to broaden the reach of data services provided by the library.

```{r top_ten_depts, echo=FALSE}
dsc_consult %>% 
  select(department) %>% 
  drop_na() %>% 
  rename(Departments = department) %>% 
  count(Departments) %>% 
  rename(Number = n) %>% 
  arrange(desc(Number)) %>% 
  head(15) %>% 
  ggplot(aes(x=reorder(Departments, -Number), y=Number)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    scale_fill_manual(values=cal_palette("kelp1")) +
    labs(x= "Department", y="Consults", title = "Consulting by Department", subtitle = "2019-2021") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    geom_text (aes(label=Number), hjust= -0.12) +
    theme_minimal()
```

```{r top_ten_depts2021, echo=FALSE}
dsc_consult %>% 
  select(department) %>% 
  drop_na() %>% 
  rename(Departments = department) %>% 
  count(Departments) %>% 
  rename(Number = n) %>% 
  arrange(desc(Number)) %>% 
  head(15) %>% 
  ggplot(aes(x=reorder(Departments, -Number), y=Number)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    scale_fill_manual(values=cal_palette("kelp1")) +
    labs(x= "Department", y="Consults", title = "Consulting by Department", subtitle = "FY 2020-21") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    geom_text (aes(label=Number), hjust= -0.12) +
    theme_minimal()
```

## Researcher Status

When provided, we also collect information on our users' status.

```{r consults_status2021, echo=FALSE}
dsc_consult %>% 
  select(ucla_affiliation) %>% 
  drop_na() %>% 
  #rename(Status = ucla_affiliation) %>% 
  count(ucla_affiliation) %>% 
  rename(Number = n) %>% 
  arrange(desc(Number)) %>% 
  ggplot(aes(x=reorder(ucla_affiliation, -Number), y=Number)) +
    geom_col(fill = "#2774AE") +
    coord_flip() +
    scale_fill_manual(values=cal_palette("kelp1")) +
    labs(x= "Status", y="Number", title="Consultations by Status: 2020-21") +
    theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1))  +
    geom_text (aes(label=Number), hjust= -0.12) +
    theme_minimal()
```

```{r}
# Source the text analysis script
source("text_analysis.R")

# Load the combined tokens
combined_tokens <- readRDS("data/combined_tokens.rds")

# View the first few rows
head(combined_tokens)
```

```{r}
library(ggplot2)

remove_bigrams <- c("data", "project", "analysis", "geospatial", "geospatial data",
                    "data gis", "wrangling", "cleaning", "discuss", "meeting", 
                    "research", "https", "time", "create", "files", "google", 
                    "data data", "collection", "cleaning data", "planning", 
                    "management", "set", "forward", "sharing", "tool", "file", 
                    "analysis coding", "management planning", "twitter", "access", 
                    "statistical", "hoping", "manipulation", "consulting")
# Remove specific words, e.g., "data"
filtered_tokens <- combined_tokens %>%
  filter(!bigram %in%  remove_bigrams)

# Combine "programming", "coding", and "coding programming" into "coding/programming"
filtered_tokens <- filtered_tokens %>%
  mutate(bigram = if_else(bigram %in% c("programming",
                                        "coding", "coding programming", "code"), 
                          "coding/programming", bigram))

# Count the frequency of each bigram
bigram_counts <- filtered_tokens %>%
  count(bigram, sort = TRUE)

# View the top bigrams
#head(bigram_counts, 20)

# Visualize the top bigrams
top_bigrams <- bigram_counts %>%
  top_n(15, n) %>%
  arrange(desc(n))

ggplot(top_bigrams, aes(x = reorder(bigram, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 15 Topics in Consulting Requests",
       x = "Bigrams",
       y = "Frequency")
```


