---
title: "What the Data Science Center supplies to `r yaml::read_yaml('divisions.yml')[[params$division]]$short`"
subtitle: "A brief for the dean, UCLA Library Data Science Center"
date: today
params:
  division: humanities
format:
  html:
    embed-resources: true
    toc: false
    css: ../styles-dean-update.css
    theme: none
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(tidyverse)
source("../src/etl/suppression.R")

div <- yaml::read_yaml("divisions.yml")[[params$division]]
stopifnot(!is.null(div))

xwalk <- read_tsv("../data/reference/instruction/department_to_organizational_parent_v2.tsv",
                  show_col_types = FALSE) |>
  select(standardized_department, organizational_parent)

# Workshop attendance records, UCLA-affiliated, Nov 2020 to Apr 2024
load("../data/processed/instruction/ucla_workshops.rda")
ws <- ucla_workshops |> left_join(xwalk, by = "standardized_department")
ws_total  <- nrow(ws)
ws_mapped <- sum(!is.na(ws$organizational_parent))
ws_div    <- ws |> filter(grepl(div$parent_pattern, organizational_parent))
ws_range  <- range(as.Date(ws$date), na.rm = TRUE)
ws_depts  <- ws_div |> count(standardized_department, name = "n", sort = TRUE) |>
  suppress_small_cells("standardized_department", "n",
                       other_label = "Other departments (each fewer than 5)")

# Consultation records by department (already suppressed at export)
cons <- read_csv("../data/processed/consultations/public/consult_records_by_department.csv",
                 show_col_types = FALSE) |>
  left_join(xwalk, by = c(department = "standardized_department"))
cons_win <- read_csv("../data/processed/consultations/public/consult_records_window.csv",
                     show_col_types = FALSE)
cons_div <- cons |> filter(grepl(div$parent_pattern, organizational_parent))

# Staff capacity (claim #17 and the home page)
cap <- read_csv("../data/processed/canonical/service_capacity_by_year.csv", show_col_types = FALSE) |>
  mutate(fte = dsc_staff_fte + contributed_staff_fte) |> filter(!is.na(fte))
fte_peak <- sprintf("%.1f", max(cap$fte))
fte_now  <- sprintf("%.1f", cap$fte[which.max(cap$year)])

hl <- read_csv("../data/processed/canonical/headline_aggregates.csv", show_col_types = FALSE)
h  <- function(k) hl |> filter(metric_id == k) |> pull(display) |> first()
mon <- function(d) format(as.Date(d), "%b %Y")
at_least <- function(n) if (n < 5) "fewer than 5" else paste("at least", comma(n))
comma <- scales::comma
```

This brief shows what DSC supplies to research in `r div$noun`, what the division relies on that DSC runs for the whole campus, and the question a reorganization raises. The research belongs to the faculty and students named; DSC's part is stated in each case.

## Research DSC has supported

```{r examples, results='asis'}
for (ex in div$examples) {
  cat("### ", ex$title, "\n\n", ex$text, "\n\n", sep = "")
  cat("*What DSC supplied:* ", ex$supplied, ". [Source](", ex$link, ")\n\n", sep = "")
}
```

## Documented use by the division

These counts cover only records whose department is known and maps to the division. Records with no department are not counted anywhere, so these are floors, not the division's full use.

- **Workshops:** `r at_least(nrow(ws_div))` attendance records from `r div$short` departments, `r mon(ws_range[1])` to `r mon(ws_range[2])`. Across UCLA, only about half of attendance records in that window have a department that maps to a school or division (838 of 1,575, claim #6).
- **Consultations:** `r at_least(sum(cons_div$records))` consultation records from `r div$short` departments, `r mon(cons_win$first_date)` to `r mon(cons_win$last_date)`. Departments with fewer than 5 records are not published, so the true figure is likely higher.

```{r depts}
if (nrow(ws_depts) > 0) {
  knitr::kable(ws_depts, col.names = c("Department (workshop attendance)", "Records"),
               format.args = list(big.mark = ","))
}
```

`r suppression_note(ws_depts, "departments")`

## What the division relies on that DSC runs for all of UCLA

These are campus services available to `r div$noun`. They are not evidence that any particular department uses them.

- **UCLA Dataverse**, the university's research-data repository: `r h("dataverse_datasets_released")` released datasets (Feb 2026).
- **Governed access to restricted data** through Redivis, including the national voter file.
- **3D capture and hosting** on pointcloud.ucla.edu.
- **Workshops and consultation** open to every UCLA researcher: `r h("instruction_sessions_lifetime")` workshop sessions since 2017, and at least `r h("people_reached_confirmed_ucla")` confirmed UCLA researchers reached campus-wide (not broken down by division).
- **DataSquad**, the student consulting team that works on faculty projects.

## The question for the division

DSC's staff time fell from `r fte_peak` FTE in 2022-23 to `r fte_now` in 2025-26. The work above does not stop needing someone when a unit changes shape. **Who will maintain access, technical support, and stewardship for these functions in `r div$noun`, and with what capacity?**

---

Figures, definitions, and sources: [Research Enabled by DSC](https://ucla-data-science-center.github.io/dsc-stats-reports/). Each figure has an entry in the site's claim register.

Tim Dennis\
Director, Data Science Center, UCLA Library
